### 2021-02-20 SupportingInformation.Rmd
Draft and analyses started!

---
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
    toc: no
    number_sections: false
    keep_tex: true
    includes:
      in_header: "sipreamble.tex"
bibliography: "references.bib"
title: "Paper title"
author:
  - Author 1^[Institution 1, email1]
  - Author 2^[Institution 2, email2]
---

```{r sisetup, echo=F}
options(tinytex.clean = FALSE)
```

# Supporting Information

\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}
\setcounter{figure}{0}
\setcounter{table}{0}

\clearpage
\newpage

# Supporting References