### 2021-03-02 Manuscript.Rmd
add supp figures and msl dfe and mnm analysis

---
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
    toc: no
    number_sections: false
    keep_tex: true
    includes:
      in_header: "preamble.tex"
    citation_package: natbib
    extra_dependencies: ["float"]
bibliography: "references.bib"
biblio-style: "genetics"
title: Linkage disequilibrium in conserved domains reveals opposing selective interactions in genes
author:
header-includes:
- |
  ```{=latex}
  \usepackage{authblk}
  \author{Aaron P. Ragsdale \\ aaronpeaceragsdale@gmail.com}
  \affil{Department of Human Genetics, McGill University, Montreal, Canada \\
         National Laboratory of Genomics for Biodiversity, Irapuato, Mexico \\
         }
  ```
date: \today
---

```{r setup, echo=FALSE}
```

\begin{abstract}

Selected mutations interfere and interact with evolutionary processes at nearby
loci, distorting patterns of allele frequencies and correlations between pairs
of mutations. A number of recent studies have used patterns of linkage
disequilibrium between selected variants to test for selective interference and
epistatic interactions, with some disagreement over interpretation of
observations. Interpreting this data is hindered by the relative lack of
analytic or even numerical expectations for patterns of variation between pairs
of loci under the combined effects of selection, dominance, epistasis, and
demography. Here, I develop a numerical approach to compute the expected
two-locus sampling distribution under diploid selection, with arbitrary
epistasis and dominance, and variable recombination and population sizes. I use
this to explore how different models of epistasis and dominance affect expected
signed LD, including for non-steady-state demography relevant to human
populations. Finally, I use whole-genome sequencing data from humans to asses
how well we can differentiate modes of selective interactions in practice. I
find that observations of positive LD between missense mutations within genes
is driven by strong positive correlations between pairs of mutations that fall
within the same conserved domain, pointing to compensatory mutations or
antagonistic epistasis as the prevailing mode of interaction within but not
outside of conserved genic elements. The heterogeneous landscape of both
strength of selection and selective interactions within protein-coding genes
calls for more refined inferences of the joint distribution of fitness and
interactive effects, and the methods presented here should prove useful in that
pursuit.

\end{abstract}

# Introduction

Most new mutations that affect fitness are deleterious and tend to be
eliminated from a population. The amount of time that a deleterious mutation
segregates depends on the strength of selection against genomes that carry it,
with very damaging mutations kept at low frequencies and purged relatively
rapidly. In the time between mutation and fixation or loss, selected variants,
both beneficial and damaging, can dramatically impact patterns of variation in
nearby linked regions. This distortion away from neutral expectations has been
widely documented in sequencing data, but questions remain about the primary
mode of interactions between multiple linked variants and their joint effects
on genome-wide patterns of diversity.

In their foundational paper, @Hill1966-gv recognized that linked selected
variants reciprocally impede the efficacy of selection at each locus. Linked
selection reduces the fixation probability of advantageous mutations and
increases that of deleterious mutations, compared to expectations under
single-locus models [@Birky1988-jm]. Allele frequency dynamics of linked
selected variants are also predicted to deviate from expectations without
interference. Under a multiplicative fitness model, where the fitness reduction
of a haplotype carrying multiple deleterious variants is equal to the product
of the fitness reduction of each mutation independently, we should expect to
see those mutation segregate on different haplotypes more often than together,
leading to negative, or repulsion, linkage disequilibrium (LD), although the
extent of LD depends non-trivially on the strength of selection and probability
of recombination separating loci [@Hill1966-gv; @McVean2000-ox].

Non-additive effects, including dominance (i.e., interactions *within* loci)
and epistasis (interactions *between* loci), further complicate our
evolutionary models. A large fraction of nonsynonymous coding mutations are
thought to be at least partially recessive [@Agrawal2011-wb; @Huber2018-cp],
with average levels of dominance correlating with strength of selection
[@Kacser1981-nc], and dominance plays an important role in shaping expected
equilibrium allele frequencies and mutation load of strongly damaging disease
mutations [@Clark1998-kq]. On the other hand, epistasis differentially impacts
the deleterious load for asexually and sexually reproducing organisms
[@Kimura1966-gc; @Kondrashov1995-va], has been invoked as an explanation for
the evolutionary advantage of sex [@Kondrashov1982-sf; @Charlesworth1990-kw;
@Barton1998-lu], and can drive incompatibilities that lead to postzygotic
isolation during the process of speciation [@Turelli2000-kz]. Within
populations, epistasis is known to cause signed LD to deviate dramatically from
zero [@Charlesworth1990-kw; @Kondrashov1995-va]. However, despite dominance
being appreciated as important factors shaping linked variation
[@Turelli2000-kz; @Zhao2016-bb] and the evolutionary importance of epistatic
interactions, we currently lack models for predicting correlation patterns
between linked mutations under general selection models.

In this paper, I develop a numerical approach to solve for the two-locus
sampling distribution under a general diploid selection model with variable
recombination and single-population size history. I use this model to describe
how epistasis and dominance shape expected patterns of signed LD, under both
steady-state and non-equilibrium demography, that have been used to test for
interference and epistasis in population genomic data. I then turn to human
sequencing data and compare patterns of LD for synonymous, missense, and
loss-of-function mutations within and between protein-coding genes and
annotated domains. I show that while synonymous and missense variants display
similar slightly positive average LD within genes, for missense mutations this
signal is driven by correlations between pairs of mutations within, but not
between, protein-coding domains, suggesting an importance for antagonistic
epistasis or a prevalence of compensatory nonsynonymous mutations within
conserved elements. Finally, I discuss how well we can expect to discriminate
between modes of selective interactions using genome-wide patterns of LD.

## Empirical observations

The most direct way to test for interactions between linked selected variants
is through deep mutation scanning experiments, in which many distinct mutations
are introduced within a target gene and then organismal fitness or some protein
function is experimentally measured [@Romero2009-yi; @Bank2015-vq;
@Puchta2016-rx; @Steinberg2016-is]. For example, using the model system of the
TEM-1 $\beta$-lactamase gene in *E. coli*, @Bershtein2006-bk found evidence for
synergistic epistasis, where multiple deleterious mutations had a greater
effect on fitness than the multiplication of the observed effects of the
individual mutations. The scale of mutation scanning experiments continues to
improve dramatically, promising greater resolution of the fitness landscape in
such model systems which can be compared to evolutionary models
[@Otwinowski2018-zk].

In most natural populations, directed mutational studies are not possible, and
we must turn to population genetic approaches to infer selective interactions
between observed mutations. Motivated by theory that linked negatively selected
mutations will display negative LD due to interference [@Hill1966-gv], and that
epistasis will drive expected LD away from zero, a number of recent studies
have used patterns of LD within classes of putatively selected variants to
infer modes of selective interactions. In a notable study from
@Callahan2011-ac, pairs of nonsynonymous mutations were found to cluster more
than expected along lineages in the *Drosophilid* species complex, and that
those clustered mutations tended to preserve the charge of the protein and were
in positive LD compared to pairs involving synonymous mutations, suggesting
that compensatory nonsynonymous variants are tolerated and maintained. More
recently, @Taverner2020-lk replicated this finding across a diverse set of
genera, showing that such epistatic interactions are important for protein
evolution.

@Sohail2017-zq observed negative LD between loss-of-function variants in
protein-coding genes (such as stop gains and losses, frameshifts, and other
nonsense mutations) in both human and fruit fly populations, which they
interpreted as evidence for widespread synergistic epistasis between these
mutations. Within the past year or two, both @Sandler2020-vk and @Garcia2020-zn
have reevaluated patterns of LD between coding variants in humans, fruit flies,
and *Capsella grandiflora*, and suggested interference and dominance may
instead be driving patterns of LD [@Garcia2020-zn] or questioned whether LD
between loss-of-function variants is significantly different from zero
[@Sandler2020-vk].

A number of factors impede our interpretation of patterns of signed LD between
coding variants. First, for strongly deleterious or loss-of-function variants,
their rarity and low frequency means that statistical measurement of LD and
other diversity measures are very noisy. Second, comparisons are based on
theory that are confined to limiting and simplistic assumptions, including
steady-state demography, simple selection and interaction models, or unlinked
loci. To generate predictions under more complex models, we rely on expensive
forward simulations. Such simulations can help build intuition and test
inference methods, but do not efficiently provide expectations for quantities
of interest across the range of relevant parameters. Analytical and numerical
methods for expected haplotype frequencies and LD under general selective
interaction models are thus crucial for interpreting patterns of variation
observed in data.

# Results

## Expected signed LD under steady-state demography

In the Methods, I expand on the moment system developed in @Ragsdale2019-nt to
compute the expected sampling distribution $\Psi_n$ of two-locus haplotypes
under a general model of selective interactions. This sampling distribution, or
two-locus haplotype frequency spectrum, stores the expected density or observed
counts of pairs of loci with each possible haplotype frequency configuration in
a sample of size $n$. In many cases it is simpler to visualize summaries (such
as $\E[D]$) or slices of the distribution (Figure \@ref(fig:fig1)) instead of
the full three-dimensional distribution. Here, I focus on low-order LD
statistics, such as $\E[D]$ and $\E[D^2]$, and their decay with recombination
distance as these are statistics that are commonly used to test for
interactions between loci. Instead of unnormalized $\E[D^2]$ and $\E[D]$, I
consider expectations for $\sigma_d^2 = \E[D^2]/\E[p(1-p)q(1-q)]$ and
$\sigma_d^1 = \E[D]/\E[p(1-p)q(1-q)]$. These statistics have two benefits:
first, the mutation rate cancels so that expectations are robust to assumptions
about the per-base mutation rate, and second, analytic expectations for these
quantities are available under neutrality at constant population size for
comparison [@Ohta1971-yd].

I first consider the case of additive selection and epistasis, followed by
dominance acting within loci but no epistatic effects, and then describe a
general diploid selection model and consider simple gene-based dominance
effects. In what follows I focus on parameters where the strengths of selection
and dominance at each locus are equal, but note that the methods and
implementation allow for arbitrary and unequal selection and dominance at the
two loci.

### Additive selection and epistasis

For additive mutations (i.e. $h=1/2$) and no epistasis, we recover the well
known Hill-Robertson interference result of negative LD between selected
mutations, which is strongest for pairs for mutations that have selection
coefficients $\gamma=O(1)$, or $s \approx 1/2N_e$ (Figure \@ref(fig:fig1)D).
For strongly deleterious mutations, LD is close to zero even with tight
linkage, as they almost always segregate at low enough frequencies that they
are unlikely to interfere with each other [@McVean2000-ox].

With epistasis, mean signed LD is large for both weakly and strongly selected
variants, with sign depending on the direction of epistatic interactions
(Figure \@ref(fig:fig2)). Synergistic epistasis results in negative LD while
antagonistic epistasis results positive LD, and large positive LD can occur
even when the epistatic effect is relatively weak. Epistasis-induced LD can
extend over long distances, especially for strongly deleterious mutations. For
example, in Figure \@ref(fig:fig1)D with $\gamma=-10$, which would be
considered moderate, large mean LD extends well beyond 100 kilobases, with more
strongly deleterious interacting mutations expected to show large signed LD
over much larger recombination distances.

### Dominance and interference

The effects of epistasis between variants on LD has been widely documented.
However, the effect of non-additive selection within a locus on interference
has received less attention. @Garcia2020-zn used large-scale forward simulation
to explore the effect of dominance on patterns of LD, showing that LD depends
on both the magnitude of the fitness and dominance effects in a nonlinear way,
but note that expensive large-scale simulations are needed to explore these
effects as no closed form approach to computing those expectations exist. The
method presented here provides such a closed form approach, so that
expectations of signed LD and other two-locus statistics can be computed more
efficiently and accurately.

The combined effect of the strength of selection and dominance on interference
is indeed nontrivial, and some parameter regimes can cause strong negative LD
between pairs of negatively selected variants. For strong selection, pairs of
recessive mutations ($h<1/2$) can have large negative signed LD (Figure
\@ref(fig:fig3)D), although it decays rapidly with increasing distance between
loci and is roughly zero for $\rho > 1$. For weakly selected mutations with
$\gamma=-1$, we see that there is no monotonic effect of the level of dominance
on negative LD, with both recessive and dominant pairs of mutations having more
negative LD than the additive case.

### General selection and gene-based dominance

While dominance and epistasis are well-characterized models of selection, a
large family of selection models can be specified by the assigning unique
selection coefficients to possible diploid pairs of haplotypes. If we assume
the diploid genotype that is homozygous for the ancestral allele ($ab/ab$) has
fitness 1, then there are nine other possible diploid two-locus genotypes that
could be given unique fitnesses (Table \@ref(tab:selmodels)), noting that
$AB/ab$ and $Ab/aB$ genotypes can have differing fitness effects.

The case with $s_{AB/ab}\not=s_{Ab/aB}$ could arise in a scenario where a
mutation at either locus within a haplotype impacts some functional region or
element, but a diploid carrying at least one copy that is free of mutations has
minimal fitness loss. This is analogous to the single-locus dominance case,
where $s_{Aa} < \frac{1}{2}s_{AA}$. In this "gene-based dominance" scenario
[e.g., @Sanjak2017-zw], an $AB/ab$ genotype has higher fitness than an $Ab/aB$
type (a simple implementation of this model is given in the final column in
Table \@ref(tab:selmodels)). Such a gene-based fitness model gives similar
expected positive signed LD to the model of antagonistic epistasis (compare
Figure \@ref(fig:fig4)E-F to G-H), although the interpretation of those two
models can vary. It does show that with a highly parameterized space of
possible general diploid selection models, multiple models with different
biological interpretations can give similar patterns of expected $D$.

## The effect of population size changes on signed LD

The moment system for $\Psi_n$ readily incorporates variable population size as
well as mutation and recombination rates and selection parameters that change
over time. Here I focus on non-equilibrium population size history and consider
scenarios relevant to human demographic history, including bottlenecks and
expansions. I explore two simple models (Figure \@ref(fig:fig4)A), one with an
instantaneous expansion and another with a bottleneck followed by recovery. I
also consider two demographic histories inferred using genome-wide genealogy
reconstruction [@Speidel2019-nj] applied to the
@1000_Genomes_Project_Consortium2015-zq dataset, and focus on size histories
for the Yoruba from Ibidan, Nigeria (YRI) and Utahns of North and West European
ancestry (CEU) (Figure \@ref(fig:fig4)B).

For each of the four histories, Figure \@ref(fig:fig4)C-H shows the dynamics of
$\sigma_d^1$ for a given parameterization of two-locus selection, including
synergistic and antagonistic epistasis, dominance within loci, and gene-based
dominance. In general across each selection model, population size expansions
are not expected to strongly affect $\sigma_d^1$, whether that expansion occurs
deeper in the past as in the simple expansion model or rapid expansion more
recently, as in the YRI. On the other hand, population size reductions push
signed LD to more extreme values and subsequent recoveries or expansion again
reduce the magnitude of LD. Under no selection condition tested here do
population size changes cause expected LD to change sign, showing that while
the magnitude of deviation of LD from zero is sensitive to population size
history, interpreting general patterns of the observed sign of LD in data
should not be strongly affected by population size history.

## Signed LD within protein-coding genes

In what follows, I turn to empirical measurements of LD within human
populations, focusing on mutations within protein-coding regions. I first look
at average values of LD across genes for synonymous, missense, and
loss-of-function mutations, and then turn to patterns of LD partitioned by
functional elements and distances separating mutations within genes.

Synonymous and missense mutations show similar levels of positive LD when
considering pairs of mutations within the same gene. Loss-of-function mutations
instead have negative LD, pointing to differing modes of selective interactions
for loss-of-function and missense mutations (Figure \@ref(fig:fig5)). Within
each population, large measurement noise gives 95\% confidence intervals that
overlap with zero for each mutation class, although the pattern is remarkably
consistent across African, European, and East Asian populations in the Thousand
Genomes dataset. Comparing mean LD within genes between African and Eurasian
populations, LD in Eurasian populations is somewhat larger on average, that
is, more positive for missense mutations and more negative for loss-of-function
mutations. This is in agreement with differences in expectations between
populations that have or have not gone through a bottleneck in their recent
past (Figure \@ref(fig:fig4)).

For pairs of missense mutations in different protein-coding genes, signed LD
decays to zero fairly rapidly and is lower than signed LD within genes even for
mutations that are more tightly linked (distance less that 50 kb). The
similarity in signed LD observed between missense and synonymous mutations
would suggest that interference between missense mutations is minimal. However,
below I show that interactive effects differ dramatically between
mutations that lie within annotated functional domains and mutations outside of
functional domains, with these opposing effects effectively canceling out when
averaging LD over entire genes. Due to the rarity of loss-of-function
mutations, I focus only on comparisons between synonymous and missense
mutations when looking at finer partitions of mutations within genes.

### Large positive LD between pairs of missense mutations in conserved domains

To further explore the observation of positive LD between both synonymous and
missense mutations within genes, I examine the role that annotated conserved
domains in protein-coding genes play in driving this signal. Note that the low
density of loss-of-function mutations within annotated domains in humans
precludes any analysis of LD between those mutations at the level of
sub-elements within genes. Protein-coding domains are conserved elements of
genes, often associated with some known functional or structural feature of a
protein. As such, we would expect that purifying selection is stronger on
average within domains than within the same gene but outside of a conserved
domains. Indeed, the site-frequency spectrum (SFS) is skewed to lower
frequencies for both missense and loss-of-function mutations within domains
when compared to the same classes of mutations outside of domains, with much
more negative values of Tajima's D within domains (Table \@ref(tab:tajimasD)).
On the other hand, no difference is observed for synonymous mutations within or
outside domains, suggesting roughly equivalent effects of selection (either
direct or linked) on synonymous variation.

For pairs of missense mutations that both fall within the same functional
domain, we observe large positive LD, roughly an order of magnitude larger than
pairs of synonymous mutations that both fall within the same domain (Figure
\@ref(fig:fig6)). This difference between pairs of missense and synonymous
variants within domains is observed for both tightly linked ($< 10$bps) and
more distant ($>1$kb) pairs (Figure \@ref(fig:lddecaydomain)). Assuming most
synonymous mutations are neutral, we would expect signed LD between missense
mutations to be less than that between synonymous mutations under predictions
from models of either Hill-Robertson interference, synergistic epistasis, or
dominance (Figures \@ref(fig:fig1)-\@ref(fig:fig3)). This observation is
opposite to those expectations, suggesting a different prevailing interactive
effect within domains than those that are commonly tested or modeled.

The large positive signal of LD for missense mutations within the same
domain does not extend to pairs of missense mutations that are in
different domains. Pairs of missense and synonymous mutations show
nearly equal levels of LD close to zero across domains, with missense mutations
even slightly more negative than synonymous mutations (Figure \@ref(fig:fig6)).
The interactive effect driving large LD within domains is therefore likely
domain-specific. However, the average distance between pairs of mutations
within domains is much smaller than between domains, so this observation may be
primarily driven by the higher recombination rate between pairs of mutations
across distinct domains.

### Reduced LD between pairs of missense mutations outside of conserved domains

I also measure LD between pairs of mutations that both fall outside of
annotated domains. For such mutations, I consider averages across all pairs
outside of domains, as well as subsets of pairs matched to the same distances
for pairs of mutation within and between domains (Figures \@ref(fig:fig6)B-D and
\@ref(fig:domainspops)). Averaging over all pairs of mutations within the same
gene but outside domains, both synonymous and missense mutations have LD
slightly above zero, with synonymous mutations observed to have slightly higher
signed LD than missense mutations. This pattern is primarily driven by tightly
linked variants. For pairs of mutations at distances matched to those within
the domains, synonymous mutations have much larger positive LD than
missense mutations, a signal that vanishes for pairs matched to distances
between domains. For larger distances, both synonymous and missense mutations
have LD approximately zero.

The role that tightly linked variants have in driving this signal can be
clearly seen in the decay of signed LD with distance (Figure
\@ref(fig:lddecay)). Both pairs of synonymous and missense mutations with
distances greater than a few hundred base pairs have LD that fluctuates around
zero. But for shorter distances, synonymous variants have large positive LD,
while missense mutations have much lower LD except for pairs within the
shortest distance bin ($<10$ bp). This reduction in LD is consistent with
Hill-Robertson interference between linked missense mutations instead of
epistatic effects, as interference between either additive or recessive
mutations without epistasis results in signed LD that decays rapidly by
recombination distance $\rho \approx 1$.

# Discussion

Previous theoretical and simulation studies have shown that interference and
interactions between selected mutations reduce the efficacy of selection at
linked loci, impacting substitution rates, the deleterious mutation load, and
dynamics of segregating mutations [@Hill1968-vu; @Birky1988-jm; @Barton1995-mj;
@McVean2000-ox]. Interference between moderately deleterious mutations and
synergistic epistasis are both expected to cause negative LD between selected
mutations, which can be readily tested using population genetic data
[@Sohail2017-zq; @Sandler2020-vk; @Garcia2020-zn]. Here, I first used a closed
numerical approach to generate expectations for signed LD under a wide range of
selective scenarios, and then compared patterns of LD in human populations
between classes of coding mutations within genes using unbiased estimators for
LD from unphased genotypes.

By taking broad genome- or gene-wide surveys of LD across functional classes of
mutations, heterogeneous patterns of interactive effects that occur within
functional sub-elements of genes can be missed. However, at the gene-wide
level, we already find large differences in interactions between pairs of
missense and loss-of-function mutations.

From gene-wide averages, LD between pairs of missense mutations does not appear
to be different from pairs of synonymous variants, but LD between
loss-of-function variants is substantially more negative. Hill-Robertson
effects under additive selection are expected to be strongest for slightly to
moderately deleterious variants (with $s\sim 1/N_e$), as strongly deleterious
mutations are not expected to interfere with one another [@McVean2000-ox].
However, loss-of-function mutations do not typically fall within this regime,
and inference of the distribution of fitness effects (DFE) for new
loss-of-function variants shows that a large majority are strongly deleterious
(Supporting Information). Instead, negative synergistic epistasis between
strongly deleterious mutations does produce large negative deviations of mean
LD. Strongly deleterious *recessive* mutations can also produce this pattern,
making it difficult to distinguish these effects with certainty. This
difficulty is compounded by the large measurement noise for $\E[D]$, especially
for loss-of-function variants for which only a few hundred pairs are found
within genes in each population. If loss-of-function variants are in fact
recessive, even partially, this can impact inferences of
the DFE (Supporting Information).

## Non-uniform interactions between selected mutations within genes

Positive average LD between missense and synonymous mutations has previously
been reported in humans and other species [@Sohail2017-zq; @Garcia2020-zn;
@Sandler2020-vk], and the similarity of their gene-wide LD might suggest that
interference or interactions between missense mutations are minimal, or at
least no stronger than those between synonymous mutations. However, averaging
over all observed pairs of mutations within a gene masks element-specific
interactive effects that drive LD in opposite directions. Nonsynonymous
mutations found within protein domains, identified as conserved subunits of a
gene with some structural or functional role, are more strongly selected
against on average, but also have increased signed LD over synonymous mutations
at the same distances within domains. Missense mutation outside of domains but
at the same distances as those within domains have more negative LD, both
compared to distance-matched synonymous mutations outside of domains and to
mutations within domains. Neither dominance effects, synergistic epistasis, nor
Hill-Robertson interference are expected to result in positive LD, so that some
other interactive effect must be driving this positive LD signal.

There are a number of possible interaction scenarios that can result in
positive signed LD between linked loci. One explanation is a
prevalence of pairs of compensatory mutations that are tolerated to
co-segregate at high frequencies within conserved domains [@Yeang2007-gj;
@Ivankov2014-tn]. @Callahan2011-ac and @Taverner2020-lk have proposed such a
mechanism to explain observed clusters of mutations in lineages of *Drosophila*
and other species. Another possibility is a model of antagonistic, or
diminishing returns epistasis, in which a single amino acid-changing mutation
within a domain damages the functionality of that subunit, but additional
mutations within that same domain reduces fitness by a factor less than the
first mutation. A third possibility, related to the antagonistic epistasis
case, is that selection acts on the functional domain as a unit instead of at
loci within the domain individually (such as gene-based dominance). In this
scenario, double heterozygotes have different fitnesses depending on whether
the mutations are found on same haplotype or on different haplotypes.

We do not see an increase in LD between mutations that are found in different
annotated domains within the same gene, which are on average at much larger
recombination distances. Nor do we find increased LD outside of domains.
Rather, for pairs of mutations outside of domains but at the same distance as
those within the same domain, missense variants have considerably lower LD than
than synonymous variants. This difference between synonymous and missense pairs
of variants largely disappears for SNPs separated by more than a few hundred or
a thousand base pairs (Figure \@ref(fig:lddecay)), which suggests that
Hill-Robertson interference is the primary mode of interaction between
mutations falling outside of domains, as epistasis is expected to impact LD
over larger distances than what is observed.

Taken together, selection on segregating variants within protein-coding genes
is nonuniform, with both the overall strength of selection and interactions
between variants differing between annotated elements. Missense and
loss-of-function mutations within conserved domains are subject to stronger
selection, skewing the SFS to lower frequencies. However, typical approaches
for inferring the distribution of fitness effects from population genomic data
average over these differences by considering all nonsynonymous mutations
together [@Boyko2008-zk; @Kim2017-xo]. It would be straightforward to adapt
those methods to infer more detailed and accurate representation of the
heterogeneous effects of new mutations within genes by partitioning by missense
and loss-of-function classes as well as by annotated domains. Similarly, the
different modes of interaction between mutations across annotated domains mean
that our standard models and simulation approaches may be too simple and will
not capture the evolutionary trajectories and patterns of diversity that differ
at finer scales.

## Clustered mutations account for positive LD among synonymous mutations

In a single randomly mating population, neutral mutations are expected to have
average signed LD of zero, but across all populations analyzed here, we observe
positive LD between synonymous mutations. While selection on some subset of
synonymous variants is possible, it should be much weaker on average than
between missense mutations, and any interference between selected synonymous
variants should lead to negative LD. @Sandler2020-vk proposed that spatial
population structure is responsible for the increase of LD for both synonymous
and missense variants, arguing that every natural population lives in some
geography that will necessarily induce some structure, and they used forward
simulations to demonstrate this effect. However, this increase in LD, at least
in simulations of spatial structure and migration, typically extends over larger
distances separating base-pairs than we observe in the data. Instead, positive
LD between synonymous mutations decays to zero within just a few hundred
base-pairs (Figures \@ref(fig:lddecay) and \@ref(fig:lddecaydomain)).

Nonrandom mutational processes provide an alternative explanation for positive
LD between neutral variants separated by short distances, in which clusters of
mutations occur simultaneously in the same mutational event. Such
multinucleotide mutations have been shown to affect patterns of LD over short
distances, on the order of 10s to 100s of base pairs [@Harris2014-zg], and
clustered mutational events may be common in humans [@Besenbacher2016-ac].
Indeed, a simple exponential model in which the fraction of pairs of mutations
arising arising from a multinucleotide event decays with distance fits the
observed decay of $\sigma_d^1$ between synonymous mutations (Supporting
Information). We may therefore treat positive LD as the baseline expectation
for tightly linked variants primarily due to clustered mutations, so that
subsequent selective processes and interactions cause LD to deviate from that
expectation, as seen for missense mutations across different sub-elements of
genes. It may also be more appropriate to compare the negative LD observed
between linked loss-of-function variants to that positive baseline expectation
instead of zero, so that they are more recessive or have stronger synergistic
epistasis than if the neutral expectation were zero.

## Challenges to distinguishing highly parameterized models of selective interactions

In exploring or inferring multi-locus selective models, the number of plausible
selection scenarios becomes quite large as we relax the strict assumptions of
additivity and multiplicative fitness effects. The inclusion of dominance
effects, epistasis, or other interactions leads to a rapid increase in the
number of parameters to consider. This makes performing forward simulations
that span the range of all such selective interaction scenarios burdensome.

Instead, closed, numerical approaches to compute expectations for two-locus
statistics under a wide range of selection models, as presented here, allow us
to explore this highly parameterized space of models far more efficiently, and
it opens the possibility for performing likelihood-based inference using signed
LD or other two-locus summaries of the data. For example, inferring the joint
distribution of dominance and selection is underpowered using the SFS. Because
signed LD is sensitive to the levels of dominance (Figure \@ref(fig:fig3)),
inferring the DFE with dominance may become feasible using the joint
distribution of allele frequencies and LD.

Using a single low-order summary of signed LD, such as $\E[D]$ or $\sigma_d^1$,
is likely insufficient to confidently discriminate modes of selective
interactions between linked mutations. Both dominance and epistasis can result
in strongly negative LD, although the expected decay of LD under those
scenarios can differ. Similarly, multiple interaction models can lead to
positive signed LD. Among all interaction models, the extent of LD and rate of
its decay also depends on the underlying distribution of selection coefficients
among a class of mutations, which are unknown for a given pair of mutations, so
that we must integrate over a distribution of fitness effects. This DFE,
however, will have been inferred under a simple set of assumptions, such as
additivity and interchangeability between sites within a gene, potentially
biasing any inference using this DFE to learn about other patterns of
selection.

Finally, exploring additional summaries of the two-locus sampling distribution
could provide additional power to distinguish between interaction models. The
method developed here allows us to compute the full sampling distribution for a
given sample size, and expectations for a large family of informative two-locus
statistics can be computed directly from this distribution. These expectations
can then be compared directly to empirical observations, which can be taken
from either phased or unphased data [@Ragsdale2020-nz]. Additional work
performing such an analysis is therefore warranted, which should provide a path
forward for distinguishing between modes of selective interactions.

# Methods

## Existing theory and numerical methods

Many well-known properties of two-locus dynamics and equilibrium LD come from
early work on the multi-locus diffusion approximation [@Kimura1955-qe;
@Hill1968-vu; @Ohta1969-ie; @Ohta1971-yd]. This includes the result that
genome-wide averages of signed LD are expected to be zero under neutrality.
Under a two-locus biallelic model, where the left locus allows alleles $A$ and
$a$ and the right locus has alleles $B$ and $b$, the standard covariance
measure of LD is defined as $D = f_{AB} - f_{A}f_{B}$ where $f_{AB}$ is the
haplotype frequency of double-derived types carrying both $A$ and $B$, and
$f_{A}$ and $f_{B}$ are the marginal frequencies of the derived alleles at each
locus. This covariance decays due to both drift and recombination at a rate
proportional to the inverse of the effective population size and the distance
separating loci: \[\E[D]_{t+1} = \left(1 - \frac{1}{2N_e(t)} - r
\right)\E[D]_t.\] While $\E[D] = 0$, the variance of $D$ is non-zero, and
@Ohta1971-yd presented their groundbreaking result that the variance of $D$
under neutrality and steady-state demography, normalized by the joint
heterozygosity of the two loci, is \begin{equation} (\#eq:ohta) \sigma_d^2 =
\frac{\E[D^2]}{\E[f_A(1-f_A)f_B(1-f_B)]}\approx\frac{5 + \rho / 2}{11 + 13\rho
/ 2 + \rho^2 / 2}, \end{equation} where $\rho = 4N_e r$.

Analytic progress beyond these results has come haltingly. In the
1980s, recursions were developed to compute the two-locus sampling distribution
under neutrality, that is, the probability of observing given counts of
two-locus haplotypes in a sample of size $n$ [@Golding1984-pu]. This approach
would continue to be developed and later form the foundation for the inference
of local recombination rates from population genetic data [@Hudson2001-sg;
@McVean2004-gj].

To include selection, however, there have been relatively few advances beyond
the Monte Carlo simulation approach taken by @Hill1966-gv, albeit we now have
access to somewhat more powerful computational resources and more sophisticated
software for performing flexible forward simulation (e.g., @Thornton2014-pn,
@Haller2019-vm). Analytic results for two-locus distributions under selection
are notoriously difficult, with a few notable flashes of progress. For example,
@McVean2007-xd considered the effect of a recent sweep on patterns of LD in
loci nearby or spanning the locus under selection, and in an impressive recent
paper, @Good2020-wj presented analytic solutions for patterns of LD between
rare mutations under additive selection with epistasis. Nonetheless, such
approaches are typically confined to steady-state demography and constrained
selection models.

Numerical methods inhabit the space in between expensive discrete simulations
and limited analytic solutions, providing a more efficient and practical method
to compute expectations of two-locus diversity measures under a wider range of
parameters and demographic scenarios. @Ragsdale2017-gg used a finite
differencing approach to numerically solve the two-locus diffusion equation
with additive selection at either locus, although that paper focused on the
applicability of two-locus statistics to demographic inference. More recently,
@Ragsdale2019-nt extended the @Hill1968-vu system for $\E[D^2]$ to compute
arbitrary moments of the distribution of $D$ for any number of populations
connected by migration and admixture. This paper also showed that such a
moments-construction can be used to solve for the two-locus sampling
distribution within a single population, though it requires a moment-closure
approximation for nonzero recombination and selection. Below, I extend this
approach to model arbitrary diploid selection, which encompasses dominance,
epistasis, and other forms of selective interactions between two loci.

## The two-locus sampling distribution with arbitrary selection

The two-locus sampling distribution is the direct analog to the single-locus
site frequency spectrum (SFS) of a given sample size. Instead of describing the
density or number of mutations with a given allele count out of $n$ samples,
the two-locus distribution $\Psi_n$ stores the density or number of pairs of
loci with observed haplotype counts, so that $\Psi_n(i, j, k)$ is the number of
pairs for which we observe $i$ copies of the $AB$ haplotype, $j$ of type $Ab$,
$k$ of type $aB$, and $n-i-j-k$ of type $ab$. The size of $\Psi_n$ grows
rapidly, with $O(n^3)$ entries, which practically limits computational
approaches to moderate sample sizes and a single population.

Under neutrality, a number of approaches exist to compute $\Psi_n$, including
the recursion due to @Golding1984-pu and @Ethier1990-tb, or more recent
numerical approaches to the two-locus coalescent [@Kamm2016-ag] or diffusion
approximation [@Ragsdale2017-gg]. Practically, selection can only be included
using the forward-in-time diffusion equation [@Kimura1955-qe; @Hill1966-gv],
where a standard approach is to first solve for the continuous distribution
$\psi$ of the density of two-locus haplotype configurations in the full
population, and then integrate $\psi$ against the multinomial sampling function
to obtain $\Psi_n$.

Alternatively, @Ragsdale2019-nt showed that there exists a system of ordinary
differential equations directly on the entries of $\Psi_n$, although that paper
focused primarily on neutral dynamics and considered a model with selection at
only one of the two loci. I briefly summarize this general approach below, but
refer readers to that previous paper for detailed derivations of the drift,
recombination, and mutation terms and the moment-closure approximation.
Instead, here I focus on generalizing the selection operator to include
epistasis, dominance, and other forms of two-locus interactions.

### Moment equation for $\Psi_n$

The system of linear ordinary differential equations for the entries of
$\Psi_n$takes the form
\begin{equation}
(\#eq:system)
{\Psi}_n^{t+1}(i, j, k; t) = 
\mathcal{D}_{N(t)}\Psi_n^t
+ \mathcal{R}_{r}\Psi_{n+1}^t
+ \mathcal{U}_{u}\Psi_n^t
+ \mathcal{S}_{s_A, s_B, \ldots, h_A, h_B, \ldots}\Psi_{n+2}^t.
\end{equation}
Here, $\mathcal{D}_{N(t)}$ is a sparse linear operator accounting for drift
with population size $N(t)$, $\mathcal{R}$ accounts for recombination
with per-generation recombination probability $r$ between the two loci,
$\mathcal{U}$ accounts for mutation, either under an infinite sites or
biallelic reversible mutation model, and $\mathcal{S}$ accounts for
selection.

The moment system for $\Psi_n$ can be derived directly from the diffusion
approximation, or it can be found through a more intuitive process of tracking
the dynamics of allelic states of a sample of size of $n$ from the full
population. We assume $n \ll N_e$, and $r$ and $s$ are $O(1/Ne)$ so that
multiple coalescence, recombination, or selective events are rare in any given
generation. For example, for drift events, we account for the probability that
two lineages within the $n$ tracked lineages at time $t+1$ coalesced in the
previous generation and, given a coalescence event, the probability that
haplotype counts changed due to this coalescence event [see Supporting
Information; @Jouganous2017-pq; @Ragsdale2019-nt]. In typical diffusion
approximation fashion, we multiple through by $2N_{ref}$ so that time
is measured in $2N_{ref}$ generations, and we consider scaled parameters
$\rho = 4Nr$, $\theta = 4Nu$, and $\gamma=2Ns$.

### Moment closure

In the absence of selection and for fully linked loci (i.e., $\rho=0$), the
system is closed and can be solved exactly. However, for nonzero recombination
or selection, the entries of $\Psi_n$ rely on the slightly larger frequency
spectra with sample sizes $n+1$ (for recombination and additive selection) or
$n+2$ (for non-additive selection). This is because if a recombination event
occurs within one of $n$ lineages being tracked by $\Psi_n$, we need to draw an
additional lineage from the full population to recombine with that chosen
lineage, thus requiring $\Psi_{n+1}^t$ to find $\Psi_n^{t+1}$. Selection events
similarly require extra lineages from the full population, which replace a
chosen lineage that fails to reproduce with probability proportional to its
selection coefficient, assuming negative selection.

This requirement of extra lineages for nonzero recombination and selection
means that the system in \@ref(eq:system) is not closed, so that we need a
moment-closure approximation to solve for $\Psi_n$. As in @Ragsdale2019-nt, a
jackknife approximation is used to estimate $\Psi_{n+l}$, for $l=1$ or $2$,
from $\Psi_n$ [@Jouganous2017-pq], so that $\hat{\Psi}_{n+l}(i, j, k) =
\mathcal{J}_{n, l}\Psi_n$, although other closure approximations are possible.
This emits a closed approximate system,
\begin{equation}
(\#eq:closed-system)
\dot{\hat\Psi}_n(i, j, k; t) = 
\mathcal{D}_{\nu(t)}\hat\Psi_n(t)
+ \mathcal{R}_{\rho}\mathcal{J}_{n, 1}\hat\Psi_{n}(t)
+ \mathcal{U}_{\theta}\hat\Psi_n(t)
+ \mathcal{S}_{\gamma, h}\mathcal{J}_{n, 2}\hat\Psi_{n}(t).
\end{equation}

The jackknife approximation, which approximates an entry $\Psi_{n+l}(i,j,k)$
using nearby entries in $\Psi_n$, is more accurate for larger sample sizes,
creating a tension between efficiency and accuracy: larger sample sizes result
in more accurate solutions, as error in the jackknife is diminished, but
computational complexity also grows rapidly in the number of entries of
$\Psi_n$, which is $O(n^3)$. Accuracy and runtime are shown in Figure
\@ref(fig:jackknife). In the results presented in this paper, sample sizes
between $n=30$ and $n=80$ are used.

Derivations for the drift, recombination, and mutation operators and the
jackknife moment-closure approximation can be found in section S1.3 of
@Ragsdale2019-nt, and I repeat the main results in the Supporting Information
of this paper.

### Selection models with epistasis and dominance

To include selection, we consider a model where we draw lineages uniformly from
the previous generation, but keep lineages with probability proportional to
their fitness. In the absence of dominance, selection reduces to a haploid
model, with acceptance and rejection probabilities depending on the fitnesses
of each haploid copy, where haplotype $Ab$ has fitness $1 + s_{A}$, $aB$ has
fitness $1 + s_{B}$, and $AB$ has fitness $1 + s_{AB}$. We assume the doubly
ancestral haplotype $ab$ has fitness $1$, so fitnesses are relative to that of
$ab$ haplotypes. The standard multiplicative fitness function assumes that
$s_{AB} \approx s_{A} + s_{B}$, and a general epistasis function can be written
as $s_{AB} = (s_{A} + s_{B}) (1 + \epsilon)$, so that $\epsilon > 0$ implies
synergistic epistasis, while $\epsilon < 0$ implies antagonistic epistasis.

To obtain the recursion equation under selection we consider drawing $n$
lineages from generation $t$, and this sampling has an expected distribution of
haplotype counts given by $\Psi_n^t$. However, assuming $s\leq0$ for each
haplotype, each of those sampled lineages as probability of being rejected
equal to the relative fitness of its haplotype state. If a lineage is rejected,
a replacement is drawn from among the population. Under the assumption that
$ns_{\cdot\cdot} \ll 1$, the probability that more than two selection events
occur in any given generation is negligibly small, so that the case of multiple
simultaneous rejections can be ignored, and $Psi_n^{t+1}$ relies only on
$\Psi_n^t$ and $\Psi_{n+1}^t$ for additive selection. The full selection
operator $\mathcal{S}$ is given in the Supporting Information.

To account for dominance within loci, or other general forms of two-locus
selection, the selection operator no longer reduces to individual haplotypes,
but instead we need to know the state of two-locus genotypes. For example, the
fitness of an individual carrying an $Ab$ haplotype depends on whether their
second haplotype is $ab$, $Ab$, $aB$, or $AB$. We can therefore assign a
selection coefficient to each possible diploid configuration, $s_{AB/AB}$,
$s_{AB/Ab}$, and so on. Assuming that the doubly homozygous ancestral $ab/ab$
genootype has relative fitness 1, this gives nine possible unique selection
coefficients in the most general two-locus selection model. Note that $AB/ab$
and $Ab/aB$ genotypes need not have the same selection coefficient, which
allows for simulation under a gene-based model of dominance (Table
\@ref(tab:selmodels)).

The general selection operator follows the same approach as the haploid
selection operator with epistasis described above. Now, in the case of a
selection event rejecting a lineage within our tracked samples, we need to draw
not only the replacement lineage from the full population but also a second
haplotype from the full population to form the diploid genotype, as this
determines the probability that we reject the focal haplotype. We again assume
that $ns \ll 1$ for all genotype selection coefficients, so that we may assume
at most a single selection event occurs in any given generation. This means
that to find $\Psi_n^{t+1}$ under a general two-locus selection model, which
encompasses dominance within either locus, gene-based dominance, or a
combination or dominance and epistatic effects, we need $\Psi_{n+2}^t$. Again,
a full derivation and expressions for the general selection operator are given
in the Supporting Information.

### Low-order summaries of the sampling distribution

From $\Psi_n$, expectations for any two-locus statistic can be found by
downsampling to the appropriate sample size. For example, to compute $\E[D]$,
the sum is taken over all haplotype configurations $\mathbf{n} = (n_{AB},
n_{Ab}, n_{aB}, n_{ab})$, weighted by the density $\Psi_n$ for that
configuration:
\begin{equation}
\E[D] = \sum_{\mathbf{n}} \Psi_n(\mathbf{n})
\frac{n_{AB}n_{ab} - n_{Ab}n_{aB}}{n(n-1)}.
\end{equation}
For large sample sizes, this is approximately equal to computing $D$ by taking
the maximum likelihood estimate for each allele frequency $f_i = n_i / n$, but
the maximum likelihood estimate will be noticeably biased for small to moderate
sample sizes. Other low-order two-locus statistics can be computed using the
same approach, as implemented in *moments* following @Ragsdale2020-nz, which
can be compared across sample sizes and between estimates from phased or
unphased data. In this paper, I focus on $\sigma_d^2 = \E[D^2] /
\E[p(1-p)q(1-q)]$ and $\sigma_d^1 = \E[D²] / \E[p(1-p)q(1-q)]$, averaged over
alleles at all frequencies. However, allele-frequency conditioned statistics
(e.g., keeping only loci below some frequency threshold [@Good2020-wj]) can be
considered using this same approach.

### Simulations of non-steady-state demography

## Analysis of human genomic data

Using the annotated variant call format (VCF) files from the phase 3
@1000_Genomes_Project_Consortium2015-zq (Thousand Genomes) data release to
subset the genotype VCFs of the same release to autosomal variants that were
annotated as either synonymous or nonsynonymous (including both missense and
more damaging mutations such as stop gains, stop losses, and so on). I further
subset to samples within each non-admixed population in the African, European,
and East Asian continental groups. Signed LD is sensitive to ancestral state
identification, so I only kept sites for which ancestral alleles were estimated
with high confidence in both the VCF info fields and the Thousand Genomes human
ancestor reconstructed from a phylogeny of six primates.

In addition to ancestral state misidentification, measured LD is sensitive to
phasing error, so I computed LD statistics using unphased genotypes following
@Ragsdale2020-nz. This approach provides unbiased estimates for pairwise LD,
under the assumption of a random mating population. I considered pairs of
mutations within the same mutation class (synonymous, missense, and
loss-of-function) either within the same gene, in different genes within
distance bins up to 500kb, or inside or outside of annotated domains within the
same protein-coding genes. I used a dataset of all known conserved domains in
humans compiled by @Stanek2020-pa.

## Data and software availability

All data and software used in this paper are publicly available and open source.
I downloaded the Thousand Genomes annotations and genotypes VCFs from the ftp
server at \url{ftp://ftp.1000genomes.ebi.ac.uk/vol1/ftp/release/20130502/}, and
the Thousand Genomes human ancestor fasta file from
\url{ftp://ftp.1000genomes.ebi.ac.uk/vol1/ftp/phase1/analysis_results/supporting/ancestral_alignments/}.
Protein domain information from @Stanek2020-pa was downloaded from
\url{http://prot2hg.com/dbdownload.php}.

Implementation of moment equations to compute expectations for two-locus and
linkage disequilibrium statistics are implemented in Python using Numpy
[@Harris2020-pc] and sparse matrix solvers in Scipy [@Virtanen2020-kr]. These
methods are packaged within *moments*, and analyses here were performed using
version 1.1.0 (\url{https://bitbucket.org/simongravel/moments/src/master/}).
Scripts to run all analyses, recreate figures, and compile this manuscript are
available at \url{https://github.com/apragsdale/two_locus_selection}.

# Acknowledgements
- Simon Gravel
- Kevin Thornton
- Alex Diaz-Papkovich
- Mashaal Sohail
- Matthias Steinrücken

\clearpage
\newpage

# Figures

```{r fig1, echo=F, fig.cap="(ref:fig1cap)", fig.align="center"}
knitr::include_graphics("figures/fig1.pdf")
```

(ref:fig1cap) \textbf{The two-locus sampling distribution and additive
selection.} (A and B) The distribution of $AB$ haplotype counts in a sample
size of 30, in which we observe 10 $A$ alleles at the left locus, and 10 $B$
alleles at the right locus. (C) The decay of normalized squared LD
($\sigma_d^2$) with recombination distance ($\rho=4N_er$) for pairs of variants
with a given selection coefficient. The dashed line is the neutral expectation
from @Ohta1971-yd, given in Equation \@ref(eq:ohta). (D) The decay of
$\sigma_d^1=\E[D]/ \E[p(1-p)q(1-q)]$ for pairs of variants. The
negative LD effect of Hill-Robertson interference is strongest for intermediate
selection, with $\gamma\sim-1$ to $-5$.

```{r fig2, echo=F, fig.cap="(ref:fig2cap)", fig.align="center"}
knitr::include_graphics("figures/fig2.pdf")
```

(ref:fig2cap) \textbf{Additive selection with epistasis.} (A, B) Synergistic
epistasis decreases $D^2$, while antagonistic epistasis increases $D^2$. (C, D)
Epistasis expected signed LD away from zero, an effect which can extend over
long recombination distances. Here, $\gamma$ is the selection coefficient for
derived alleles $A$ and $B$, assuming selection is equal at the two loci.
$\epsilon=0$ is the case of pure additive selection with no epistasis, and
$\gamma_{AB}=(\gamma_A + \gamma_B)(1+\epsilon)$. Dashed lines are expectations
under neutrality for both $\sigma_d^2$ and $\sigma_d^1$.

```{r fig3, echo=F, fig.cap="(ref:fig3cap)", fig.align="center"}
knitr::include_graphics("figures/fig3.pdf")
```

(ref:fig3cap) **Dominance effects.** (A-B) Pairs of recessive mutations
($h<0.5$) have higher squared LD than additive mutations, while pairs of
dominant alleles have lower expected $D^2$. (C-D) Pairs of recessive, strongly
deleterious mutations can have large negative LD, although this correlation
decays more rapidly with larger recombination than LD under epistasis (Figure
\@ref(fig:fig2)D). Notably, there is no simple relationship between strength of
selection and dominance, particularly when $s$ is $O(1/N)$.

```{r fig4, echo=F, fig.cap="(ref:fig4cap)", fig.align="center"}
knitr::include_graphics("figures/fig4.pdf")
```

(ref:fig4cap) **The effects of demography on signed LD.** Left column:
simulations under two toy models, one with an instantaneous expansion in the
past, and the other with a 10-fold reduction in size followed by recovery.
Right column: simulations under population size histories inferred using Relate
[@Speidel2019-nj] applied to populations from
@1000_Genomes_Project_Consortium2015-zq. In all simulations, $\gamma$ is set to
$-2$ at both loci. (A-B) Model illustrations. Note that the differences in
scale. (C-D) Additive selection with no epistasis, showing fully linked
($\rho=0$) and more loosely linked ($\rho=5$) pairs. Bottleneck events result
in larger negative values of signed LD, but population size changes are not
expected to change the sign of mean LD. (E-F) Synergistic ($\epsilon > 0$) and
antagonistic ($\epsilon < 0$) epistasis. Again, bottleneck events result in
larger signed LD. Here, $\epsilon=1/2$ is a model with $\gamma_A=-2$,
$\gamma_B=-2$ and $\gamma_{AB}=-6$, and $\epsilon=-1/2$ is a model with
$\gamma_A=-2$, $\gamma_B=-2$ and $\gamma_{AB}=-2$. (G-H) Dominance within loci
results in negative LD, while dominance acting at the gene level results in
positive LD. Again, bottlenecks result in larger average LD in both cases.

```{r fig5, echo=F, fig.cap="(ref:fig5cap)", fig.align="center"}
knitr::include_graphics("figures/fig5.pdf")
```

(ref:fig5cap) **LD within and between protein-coding genes.** (A-B) Pairs of
variants in coding regions partitioned by classes of mutations.
Loss-of-function mutations are defined as stop and start gains and losses,
frameshifts, and other nonsense mutations. In all populations, we observe
positive mean signed LD, although the confidence interval of each measurement
overlaps with zero. Similarly, we observe mean negative LD for pairs of
loss-of-function variants within genes. There are only $\sim200-300$ pairs of
loss of function variants that fall within the same gene across the populations
tested here, causing large uncertainty in each measurement. Gray lines indicate
2 standard errors of the measurement, via bootstrap. (C-F) While pairs of
synonymous and missense mutations within genes are consistently positive, pairs
between genes are not consistent in their sign across populations. LD for all
populations analyzed here are shown in Figure \@ref(fig:genespops).

```{r fig6, echo=F, fig.cap="(ref:fig6cap)", fig.align="center"}
knitr::include_graphics("figures/fig6.pdf")
```

(ref:fig6cap) **LD within and between annotated domains in protein-coding
genes.** Considering only variants that fall within annotated domains in
protein-coding genes, pairs of mutations that fall within the same gene display
larger positive LD than pairs that lie in different domains. This pattern is
most striking for missense mutations, for which pairs within domains have
average signed LD that is consistently larger than for pairs that span two
domains, or for pairs of synonymous mutations that are within the same domain.
The slightly positive LD that we observe for missense and synonymous mutations
gene-wide (Figure \@ref(fig:fig4)A) is largely driven by the subset of pairs
mutations that fall within the same annotated domain. LD for all populations
analyzed here are shown in Figure \@ref(fig:domainspops).

\clearpage
\newpage

# References