### 2022-06-03 manuscript/paper.tex
updates for minor revisions

\documentclass[]{article}
\usepackage[round]{natbib}

\usepackage[margin=1in]{geometry}
\usepackage{url} 
\usepackage{authblk}
\usepackage{graphicx}
\usepackage{color}
\usepackage{booktabs}
\usepackage{xspace}

\usepackage{amsmath,amssymb}
\usepackage{longtable,booktabs,array}

\newcommand{\E}{\mathbb{E}}
\newcommand{\moments}{\texttt{moments}\xspace}
\newcommand{\Relate}{\texttt{Relate}\xspace}
\newcommand{\msprime}{\texttt{msprime}\xspace}

\newcommand{\comment}[1]{{\textcolor{red}{APR: #1}}}

% cross-reference with supplement
\usepackage{xr}
\externaldocument{supplement}

\usepackage{lineno}
%\linenumbers

\begin{document}

%\title{A heterogeneous landscape of selection and epistasis in protein-coding genes revealed by two-locus statistics}
\title{Local fitness and epistatic effects lead to distinct patterns of linkage disequilibrium in protein-coding genes}
\author[]{Aaron P. Ragsdale}
\affil[]{Department of Integrative Biology, University of Wisconsin--Madison, WI, USA}
\affil[]{apragsdale@wisc.edu}
\maketitle

\begin{abstract}

Selected mutations interfere and interact with evolutionary processes at nearby
loci, distorting allele frequency trajectories and creating correlations
between pairs of mutations. Recent studies have used patterns of linkage
disequilibrium (LD) between selected variants to test for selective
interference and epistatic interactions, with some disagreement over
interpreting observations from data. Interpretation is hindered by a lack of
analytic or even numerical expectations for patterns of variation between pairs
of loci under the combined effects of selection, dominance, epistasis, and
demography. Here, I develop a numerical approach to compute the expected
two-locus sampling distribution under diploid selection with arbitrary
epistasis and dominance, recombination, and variable population size. I use
this to explore how epistasis and dominance affect expected signed LD,
including for non-steady-state demography relevant to human populations. Using
whole-genome sequencing data from humans, I explore genome-wide patterns of LD
within protein-coding genes. I show that positive LD between missense mutations
within genes is driven by strong positive allele-frequency correlations between
mutations that fall within the same annotated conserved domain, pointing to
compensatory mutations or antagonistic epistasis as the prevailing mode of
interaction within conserved genic elements. LD between missense mutations is
reduced outside of conserved domains, as expected under Hill-Robertson
interference. This variation in both mutational fitness effects and selective
interactions within protein-coding genes calls for more refined inferences of
the joint distribution of fitness and interactive effects, and the methods
presented here should prove useful in that pursuit.

\end{abstract}

\section{Introduction}\label{sec:introduction}

Most new mutations that affect fitness are deleterious and will be eliminated
from a population. The amount of time that a deleterious mutation segregates
depends on the strength of selection against genomes that carry it, with very
damaging mutations kept at low frequencies and purged relatively rapidly. In
the time between mutation and fixation or loss, selected variants can
dramatically impact patterns of variation in nearby regions
\citep{Smith1974-em,Charlesworth1995-dq,Kim2000-on}. This distortion
away from neutral expectations has been empirically documented using sequencing
data from an ever-growing set of study systems
\citep[e.g.,][]{Novembre2009-kc,Cutter2013-mm,Comeron2014-oy}, but questions remain
about the primary mode of interactions between selected variants and their
joint effects on genome-wide patterns of diversity.

In their foundational paper, \citet{Hill1966-gv} recognized that linked
selected variants reciprocally impede the efficacy of selection at each locus,
a process known as selective interference. Linked selection reduces the
fixation probability of advantageous mutations and increases that of
deleterious mutations compared to expectations under single-locus models
\citep{Birky1988-jm}. Allele frequency dynamics and correlations of linked
selected variants are also predicted to deviate from single-locus expectations.
Under a multiplicative fitness model, where the fitness reduction of a genome
carrying multiple deleterious variants is equal to the product of the fitness
reduction of each independent mutation, net linkage disequilibrium (LD) is
expected to be zero for unlinked sites \citep{Kondrashov1995-va}.  However,
mutations at linked loci are expected to segregate on different haplotypes more
often than together, leading to negative, or repulsion, LD, and the magnitude
of LD depends non-trivially on the strength of selection and the probability of
recombination separating loci \citep{Hill1966-gv,McVean2000-ox}.

Non-additive effects, including dominance (i.e., interactions \emph{within} a
locus) and epistasis (interactions \emph{between} loci), further complicate our
evolutionary models. A large fraction of nonsynonymous coding mutations are
thought to be at least partially recessive \citep{Agrawal2011-wb,
Huber2018-cp}, with average levels of dominance correlating with strength of
selection \citep{Kacser1981-nc}, and dominance plays an important role in
shaping expected equilibrium allele frequencies and the mutation load of
strongly damaging disease mutations \citep{Clark1998-kq}. On the other hand,
epistasis differentially impacts the deleterious load in asexually and sexually
reproducing organisms \citep{Kimura1966-gc,Kondrashov1995-va}, has been invoked
as an explanation for the evolutionary advantage of sex
\citep{Kondrashov1982-sf,Charlesworth1990-kw,Barton1998-lu}, and can drive
incompatibilities that lead to postzygotic isolation during the process of
speciation \citep{Turelli2000-kz}. Within populations, epistasis is known to
cause signed LD to deviate dramatically from zero \citep{Charlesworth1990-kw,
Kondrashov1995-va}. However, despite appreciation of the effect of dominance on
linked variation \citep{Turelli2000-kz, Zhao2016-bb} and the evolutionary
importance of epistatic interactions, we currently lack models for predicting
patterns of correlations between linked mutations under general selection
models.

Here I develop a numerical approach to solve for the two-locus
sampling distribution under a general diploid selection model with variable
recombination and single-population size history. I use this model to describe
how epistasis and dominance shape expected patterns of signed LD, under both
steady-state and non-equilibrium demography, that have been used to test for
interference and epistasis in population genomic data. I then turn to human
sequencing data and compare patterns of LD for synonymous, missense, and
loss-of-function mutations in protein-coding genes and annotated conserved
domains. I show that while synonymous and missense variants display similar
slightly positive average LD within genes, for missense mutations this signal
is driven by correlations between pairs of mutations within, but not between or
outside of, conserved annotated elements. This suggests an importance for
antagonistic epistasis or a prevalence of compensatory nonsynonymous mutations
within conserved gene domains.

\subsection{Empirical observations}\label{sec:empirical-observations}

The most direct way to test for interactions between linked selected variants
is through deep mutation scanning experiments, in which many distinct mutations
are introduced within a target gene and then organismal fitness or protein
function is experimentally measured \citep{Romero2009-yi, Bank2015-vq,
Puchta2016-rx, Steinberg2016-is}. Using the model system of the
TEM-1 \(\beta\)-lactamase gene in \emph{E. coli}, \citet{Bershtein2006-bk}
found evidence for synergistic epistasis, in which the combined effect of
multiple deleterious mutations on individual fitness was larger than would be
expected from multiplying the independent observed effects of each individual
mutation. The scale of mutation scanning experiments continues to increase,
promising greater resolution of the fitness landscape in such model systems
that can be compared to evolutionary theory \citep{Otwinowski2018-zk}.

Directed mutational studies are not possible in most natural populations, and
we must turn to population genetic approaches to infer selective interactions
between observed segregating polymorphisms. Motivated by theoretical
predictions that linked negatively selected mutations will display negative LD
due to interference \citep{Hill1966-gv} and that epistasis will drive expected
LD away from zero, multiple recent studies have used patterns of LD within
classes of putatively selected variants to infer modes of selective
interactions. \citet{Callahan2011-ac} observed that pairs of tightly linked
nonsynonymous mutations cluster together more than expected along lineages in
the \emph{Drosophilid} species complex, and that those clustered mutations tend
to preserve the charge of the protein and were in positive LD compared to pairs
of synonymous mutations at the same distance. From this, they proposed that
compensatory nonsynonymous variants regularly arise and are maintained. More
recently, \citet{Taverner2020-lk} replicated this finding across a diverse set
of genera, showing that such epistatic interactions are important in protein
evolution.

\citet{Sohail2017-zq} observed negative LD between loss-of-function variants in
protein-coding genes (loss-of-function mutations include stop gains and losses,
frameshifts, and other nonsense mutations) in both human and fruit fly
populations. This was interpreted as evidence for widespread synergistic
epistasis between these mutations, in which the fitness reduction of multiple
mutations is greater than the product of that of each individual mutation
independently. Both \citet{Sandler2021-of} and \citet{Garcia2021-zn}
recently reevaluated patterns of LD between coding variants in humans, fruit
flies, and \emph{Capsella grandiflora}, suggesting interference and
dominance could be driving patterns of LD \citep{Garcia2021-zn} and
questioning whether LD between loss-of-function variants is significantly
different from zero \citep{Sandler2021-of}.

Many factors impede our interpretation of patterns of signed LD between
coding variants. First, for strongly deleterious or loss-of-function mutations,
their low allele frequencies mean that measurements of LD and other diversity
statistics are very noisy. Second, comparisons are based on theory with
limiting assumptions, such as steady-state demography, simple selection and
interaction models, or unlinked loci. To generate predictions under more
complex models, we rely on expensive forward simulations. Such simulations can
help build intuition and be used to test inference methods, but they do not
efficiently provide expectations for quantities of interest across the range of
relevant parameters. Analytical and numerical methods for expected haplotype
frequencies and LD under general selective interaction models are thus crucial
for interpreting patterns of variation observed in data.

\section{Results}\label{sec:results}

\subsection{Expected signed LD under steady-state demography}
\label{sec:signed-ld}

\begin{figure}[tb!]
    %\internallinenumbers
    \centering
    \includegraphics{../figures/two-locus-spectrum/stats_summaries}
    \caption{
        \textbf{Sampling distributions and their summaries.}
        Low-order summaries of sampling distributions are commonly computed
        for allele frequencies (A, the site-frequency spectrum, SFS)
        and two-locus haplotype distributions (B, linkage disequilibrium, LD).
        Demographic and selective processes affect both the SFS and LD,
        and observations of nonzero values of Tajima's $D$ or signed LD
        ($\sigma_d^1$) are often taken as evidence for selection or interactions
        between loci, respectively.
        (C) The full two-locus haplotype sampling distribution is a three-dimensional
        object, making it difficult to visualize. We can instead visualize conditional
        distributions of the full sampling distribution, e.g., conditioned on observing
        \(n_A\) copies of the \(A\) allele at the left locus and \(n_B\) copies of
        \(B\) at the right locus \citep[e.g.,][]{Hudson2001-sg}.
        (D) $\sigma_d^2$, which is closely related to $r^2$, decays
        with increasing recombination distance between loci \citep{Ohta1969-ie}.
        Selection distorts squared LD away from neutral expectations.
        (E) $\sigma_d^1$ (signed LD) is zero for
        pairs of neutral mutations \citep{Hill1968-vu}. Interference between linked
        selected mutation causes negative signed LD \citep{Hill1966-gv}, and other
        forms of interactions between selected mutations can cause large negative
        or positive signed LD.
    }
    \label{fig:TwoLocusFS}
\end{figure}

In the Methods, I expand on the moment system developed in
\citet{Ragsdale2019-nt} to compute the expected sampling distribution of
two-locus haplotypes (\(\Psi_n\), Figure~\ref{fig:TwoLocusFS}) under a general
model of selective interactions. This sampling distribution stores the expected
density or observed counts of pairs of biallelic loci with each possible
haplotype configuration in a sample of size \(n\). Below, we compute
expectations for \(\Psi_n\) under varying scenarios of selection and
interaction between pairs of mutations. It is not possible in this framework to
include the effects of additional linked selected mutations, such as background
selection, and individual-based forward simulations are still needed for such
scenarios (e.g., Figures~\ref{fig:bgs1}--\ref{fig:bgs4}).

In many cases it is simpler to visualize summaries such as the expectation or
variance of \(D\) (Figure~\ref{fig:TwoLocusFS}D, E) or conditional slices of
the distribution (Figure~\ref{fig:TwoLocusFS}C) instead of the full
three-dimensional distribution \(\Psi_n\). For pairs of biallelic loci, with
alleles labeled \(A/a\) at the left locus and \(B/b\) at the right locus,
\(D=f_{AB}-f_A f_B\) is the standard covariance measure of LD, where \(f_{AB}\)
is the frequency of haplotypes carrying both \(A\) and \(B\), and \(f_A\) and
\(f_B\) are the marginal allele frequencies of \(A\) and \(B\). Here I focus on
low-order LD statistics and their decay with recombination distance, as these
are statistics that are commonly used to test for interactions between loci.
Instead of \(\E[D^2]\) and \(\E[D]\) I consider expectations for \(\sigma_d^2 =
\E[D^2]/\E[f_A(1-f_A)f_B(1-f_B)]\) and \(\sigma_d^1 =
\E[D]/\E[f_A(1-f_A)f_B(1-f_B)]\). Normalized statistics have two benefits: (1)
the mutation rate cancels so that expectations are robust to assumptions about
the per-base mutation rate, and (2) we can compare to analytic expectations for
these quantities under neutrality and constant population size
\citep{Ohta1971-yd}.

Below, I first consider the case of additive selection, Hill-Robertson
interference, and epistasis. I then explore the effect of dominance acting
within loci but without epistatis, and then describe a general diploid
selection model and consider gene-based dominance effects. I present results
for equal strengths of selection and dominance at each locus, but note that the
methods presented here allow for arbitrary and unequal selection and dominance
at the two loci. I also focus primarily on weak to moderate negative selection
(\(|2Ns| \approx 1 - 20\)), since this range of selection leads to the
strongest signals of interference (Figure~\ref{fig:HillRobertson}) and is the
parameter regime for which the numerical approach is most accurate.

\subsubsection{Additive selection and epistasis}
\label{sec:additive-selection}

\begin{figure}[tb!]
    %\internallinenumbers
    \centering
    \includegraphics{../figures/hill_robertson}
    \caption{
        \textbf{Hill-Robertson interference.}
        Interference between pairs of selected mutations causes negative
        signed LD.
        (A) The expected normalized variance of $D$ (\(\sigma_d^2\))
        decreases below neutral expectations
        as the strength of negative selection increases.
        (B, C) For tightly linked loci (\(4Nr \lesssim 1\)), interference is most
        noticeable for pairs of mutations with \(s \approx 1/N\).
        At larger recombination distances (\(4Nr > 1\)), signed LD is most
        negative for somewhat stronger selection coefficients.
        Dashed lines show neutral expectations.
    }
    \label{fig:HillRobertson}
\end{figure}

\begin{figure}[tb!]
    %\internallinenumbers
    \centering
    \includegraphics{../figures/epistasis_prediction}
    \caption{
        \textbf{Additive selection and epistasis.}
        Left panels (A and C) show expectations for the decay of
        \(\sigma_d^2\) with recombination
        distance, and right panels (B and D) show expectations for the
        decay of \(\sigma_d^1\). Dashed lines
        show neutral expectations.
        For both weak (\(s=-1/2N\), A and B) and moderate
        (\(s=-10/2N\), C and D) selection,
        antagonistic epistasis (\(\epsilon < 0\))
        causes positive signed LD and increased $\sigma_d^2$ over
        a multiplicative model (\(\epsilon = 0\)),
        and synergistic epistasis (\(\epsilon > 0\)) results in negative signed
        LD beyond that of Hill-Robertson interference and decreased \(\sigma_d^2\).
    }
    \label{fig:epistasis}
\end{figure}

For mutations under additive selection (\(h=1/2\)) and no epistasis, we
recover the well known \citet{Hill1966-gv} interference result of negative LD
between selected mutations, which is strongest for pairs of mutations that have
selection coefficients \(\gamma = 2N_es = O(1)\), or \(s \approx 1/2N_e\) (Figure
\ref{fig:HillRobertson}).
For strongly deleterious mutations, LD is close to zero even
with tight linkage, as they almost always segregate at low enough frequencies
that they are unlikely to interfere with each other \citep{McVean2000-ox}.

With epistasis, mean signed LD is large for both weakly and strongly selected
variants, with sign depending on the direction of epistatic interactions
(Figure \ref{fig:epistasis}).
Synergistic epistasis (in which the effect of two mutations together is larger than
the product of each individual mutation's effect) results in negative LD while
antagonistic epistasis (in which the combined effect is less than the product of
independent effects) results in positive LD, and large nonzero LD can occur
even when epistasis is relatively weak. Epistasis-induced LD can
extend over long distances, especially for strongly deleterious mutations.
Even moderately deleterious mutations with
population-size-scaled selection coefficients of \(\gamma=-10\) show large mean
LD that extends to values of \(\rho\) much greater than 1
(Figure \ref{fig:epistasis}F; for humans, assuming
roughly 1 cM/Mb, this is on the order of 100 kilobases or more).
More strongly deleterious interacting mutations are expected to show large
signed LD over much larger recombination distances.

\subsubsection{Dominance}\label{sec:dominance}

\begin{figure}[tb!]
    %\internallinenumbers
    \centering
    \includegraphics{../figures/dominance_roze}
    \caption{
        \textbf{The effect of dominance on LD.}
        (A, B) The strengths of selection and dominance interact in a nonlinear way
        to shape expected signed LD. For weakly to moderately selected mutations,
        as shown here, signed LD can be large and negative for tightly linked loci
        (e.g., \(\gamma=-5\), \(h=0\), and \(\rho < 1\)). However, this large signed
        LD decays with recombination distance faster under a model of recessivity
        than does signed LD under a model of additive selection and
        epistasis (Figure~\ref{fig:epistasis}).
        (C) Interference effects are most pronounced for recessive deleterious
        variants.
        (D, E) Recessive strongly deleterious mutations can have positive signed
        LD, as recently shown by \citet{Roze2021-cf}.
        However the dominance threshold at which \(\sigma_d^1\) switches from
        positive to negative depends on the strength of selection, and weakly
        selected mutations can show non-monotonic behavior as $h$ varies.
        Selection parameters of \(sh=-0.1\) imply extremely strong selection
        (\(h=0.05\) results in \(s=-0.2\) and \(\gamma=-400\) for homozygous
        diploids at a single locus). The numerical approach for \(\Psi_n\)
        cannot handle such strong selection.
    }
    \label{fig:dominance}
\end{figure}

The effect of non-additive selection on correlations between mutations has
received increased attention recently. For example, \citet{Garcia2021-zn} used
large-scale forward simulations to explore how dominance impacts patterns of
LD, showing that LD depends nonlinearly on the magnitudes of both selection
and dominance. \citet{Roze2021-cf} found an analytic expression
for LD between pairs of strongly deleterious mutations under steady-state
demography, showing that LD can be either positive or negative depending on the
strength of dominance.

The combined effect of the strengths of selection and dominance on interference
is indeed nontrivial (Figure~\ref{fig:dominance}A). Some parameter regimes can
cause large negative LD between negatively selected variants, with
moderately selected recessive variants showing stronger signals of interference
than additive selection (Figure~\ref{fig:dominance}B). Unlike epistatic
interactions, signed LD decays rapidly with increasing distance between loci
and is roughly zero for \(\rho \gg 1\). For weakly selected mutations
(\(|\gamma|\approx 1\)), there is no monotonic effect of the level of dominance
on negative LD, with both recessive and dominant pairs of mutations having more
negative LD than additive mutations.

Discrete simulations and the moment approach confirm the result from
\citet{Roze2021-cf}, that positive LD can occur for strong negative selection
and small values of \(h\) (Figure~\ref{fig:dominance}D, E). However, while the
analytic formula in \citet{Roze2021-cf} predicts that LD should be positive for
\(h<0.25\) and negative for \(h>0.25\), this appears to only hold in the limit
of strong selection (compare to Figure 1A in \citet{Roze2021-cf}). For moderate
to moderately strong selection, this threshold of \(h\) can be less than
\(0.25\), and LD is negative for all \(0\leq h \leq 1\) for weakly deleterious
mutations (Figure~\ref{fig:dominance}C).

\subsubsection{Arbitrary two-locus selection models}
\label{sec:general-selection}

\begin{figure}[tb!]
    %\internallinenumbers
    \centering
    \includegraphics{../figures/gene-based-compensatory}
    \caption{
        \textbf{Multiple modes of interactions can lead to large positive signed LD.}
        Both antagonistic epistasis (A and C, and which includes compensatory mutation
        models) and gene-based dominance (B and D) lead to large positive signed LD.
        Compensatory mutations (\(\epsilon \approx -1\)) also cause increased
        \(\sigma_d^2\) compared to neutral expectations (dashed black lines),
        while weaker antagonistic epistasis does not increase \(\sigma_d^2\) above
        neutral expectations (compare to Figure~\ref{fig:epistasis}C).
        Gene-based dominance instead causes lower \(\sigma_d^2\) than neutral
        expectations.
        While signed LD may be similar between different interaction models,
        other two-locus summaries of the data may help to distinguish between
        interaction models.
        }
    \label{fig:gene-based-compensatory}
\end{figure}

Beyond standard models of epistasis and dominance, a large family of selection
models can be specified by assigning unique fitness effects to each possible
diploid pair of haplotypes. Assuming the diploid genotype homozygous for the
ancestral alleles (\(ab/ab\)) has relative fitness 1, there are nine other
diploid two-locus genotypes that could be given unique fitnesses (Table
\ref{tab:selmodels}), noting that \(AB/ab\) and \(Ab/aB\) genotypes can have
differing selection coefficients.

The case with \(s_{AB/ab} \not= s_{Ab/aB}\) may arise in a scenario where a
mutation at either locus within a haplotype impacts some functional element,
but a diploid individual carrying at least one copy that is free of mutations
has minimal fitness loss. In this ``gene-based dominance'' scenario
\citep[e.g.,][]{Sanjak2017-zw}, an \(AB/ab\) genotype has higher fitness than
an \(Ab/aB\) type (Table~\ref{tab:selmodels}). Such a gene-based fitness model gives
positive signed LD, similar to the model of antagonistic epistasis
(Figure~\ref{fig:gene-based-compensatory}), although the interpretation of
those two models can differ. With a highly parameterized space of possible
general diploid selection models, multiple models with different biological
interpretations can give similar patterns of expected signed LD.

\subsection{The effect of population size changes on signed LD}
\label{sec:demography}

\begin{figure}[tb!]
    %\internallinenumbers
    \centering
    \includegraphics{../figures/demog_bottle_expand}
    \caption{
        \textbf{The effects of demography on signed LD.}
        (A) Simulations under histories of an instantaneous size expansion
        and a five-fold reduction and recovery.
        For two selection strengths (\(\gamma=-1\) and \(\gamma=-10\)) and three
        cases of interactions (non-epistatic interference (\(\epsilon=0\), B),
        synergistic epistasis (\(\epsilon=1/2\), C), and antagonistic epistasis
        (\(\epsilon=-1/2\), D), each with \(\rho=0\)), a
        sudden decrease in population size can cause
        large changes in signed LD, often in the opposite direction than more
        subtle shifts due to instantaneous expansion events.
        (E) Signed LD between pairs of rare alleles (both \(n_A, n_B \leq 4\), with
        sample size \(n=50\)) is more sensitive to population size changes.
        (F) Signed LD between pairs of common alleles (both \(n_A, n_B \geq 5\))
        is comparatively more stable over time.
        Colors are matched to conditions shown in B--D.
        Additional comparisons, including for dominance models and showing
        \(\sigma_d^2\), are shown in Figures~\ref{fig:toy_sd2}--\ref{fig:toy_dom_sd2}.
        Dashed lines indicate neutral expectations.
    }
    \label{fig:toy}
\end{figure}

\begin{figure}[tb!]
    %\internallinenumbers
    \centering
    \includegraphics{../figures/demog_YRI_CEU.dominance}
    \caption{
        \textbf{Signed LD under inferred models of human population-size history.}
        Piecewise-constant population size histories inferred by \texttt{Relate} applied to
        \citet{1000_Genomes_Project_Consortium2015-zq} phase 3 data were used to
        simulate time series of two-locus statistics, as in Figure~\ref{fig:toy}.
        (A) The CEU are inferred to have a stronger bottleneck than the YRI
        10--100ka, reflecting the out-of-Africa event.
        For fully linked loci (\(\rho=0\)) and \(\gamma=-2\) at both loci,
        I compare the effects of
        (B) standard interference, (C) site-wise dominance, and (D) gene-based
        dominance on \(\sigma_d^1\).
        As with epistasis, more severe bottlenecks have larger effects on signed LD.
        (E, F) LD among common variants is more stable than among pairs of
        uncommon variants. Colors are matched to conditions shown in B--D.
        Additional comparisons with epistasis and showing \(\sigma_d^2\) are
        in Figures~\ref{fig:relate}--\ref{fig:relate_dom_sd2}.
        Dashed lines indicate neutral expectations.
    }
    \label{fig:relate_dom}
\end{figure}

The moment system for \(\Psi_n\) readily incorporates variable population size.
I explore two simple size history models (Figure \ref{fig:toy}A), one with an
instantaneous expansion and another with a bottleneck followed by recovery. I
also consider two demographic histories inferred using genome-wide gene
genealogy reconstruction \citep{Speidel2019-nj} applied to the
\citet{1000_Genomes_Project_Consortium2015-zq} dataset, and focus on size
histories for the Yoruba from Ibidan, Nigeria (YRI) and Utahns of North and
West European ancestry (CEU) (Figure \ref{fig:relate_dom}A).

For each of the four histories,
Figures~\ref{fig:toy},~\ref{fig:relate_dom},~and~\ref{fig:toy_sd2}--\ref{fig:relate_dom_sd2}
show the dynamics of \(\sigma_d^1\) and \(\sigma_d^2\) for a given
parameterization of two-locus selection, including synergistic and antagonistic
epistasis, dominance within loci, and gene-based dominance. In general across
each selection model, population size expansions do not strongly affect
\(\sigma_d^1\), whether that expansion occurs deeper in the past as in the
simple expansion model or rapid expansion more recently, as for YRI. On the
other hand, population size reductions tend to push signed LD to more extreme
values and subsequent recoveries or expansion again reduce the magnitude of LD.
Under no selection condition tested here do population size changes cause
expected LD to change sign, showing that while the magnitude of deviation of LD
from zero is sensitive to population size history, interpretions of the
observed sign of LD should be robust to population size history.

\subsection{Signed LD within protein-coding genes}

\begin{figure}[tb!]
    %\internallinenumbers
    \centering
    \includegraphics{../figures/data_compact}
    \caption{
        \textbf{LD in human protein-coding genes and annotated domains.}
        (A) Gene-wide averages of signed LD are slightly positive
        for both missense and synonymous mutations, considering pairs of mutations
        at matching distances. This positive, equal \(\sigma_d^1\) is also observed
        when conditioning on allele frequencies or considering only common variants
        with minor allele frequencies \(\gtrsim 0.1\)
        (Tables~\ref{tab:afrsd1}--\ref{tab:eassd2}).
        (B) While there are relatively fewer pairs of loss-of-function mutations
        within genes, causing larger measurement uncertainty, they have
        negative average LD. Measurement noises for each class of
        mutations overlap with zero and with each other, making it difficult to
        draw firm conclusions on the patterns of interactions occurring gene-wide.
        (C) Partitioning pairs of mutations as falling within or outside
        of conserved domains reveals opposing patterns of signed LD, with
        \(\sigma_d^1\) between missense mutations larger than that of synonymous
        mutations within domains. Outside of conserved domains, missense mutations
        have reduced LD compared to synonymous mutations. Distances of pairs outside
        of domains were matched to within-domain mutation pair distances.
        (D--F) Rare and uncommon variants (with frequencies $<0.1$ have positive
        LD, with synonymous LD exceeding missense LD in many populations. However,
        common missense variants within domains have large positive LD and are
        responsible for the pattern seen in (C).
    }
    \label{fig:data}
\end{figure}

Here, I examine patterns of signed LD between mutations in human protein-coding
genes partitioned by functional annotations. Synonymous and missense mutations
show similar levels of slightly positive signed LD when considering pairs of
mutations within the same gene averaged over all autosomal chromosomes.
Loss-of-function mutations have more negative LD, possibly due to differing
modes of selective interactions for loss-of-function and missense mutations
(Figure~\ref{fig:data}A~and~B). Within each population, measurement noise gives
95\% confidence intervals that overlap with zero in each mutation class,
although the observed patterns are remarkably consistent across African,
European, and East Asian populations in the Thousand Genomes dataset. Comparing
mean LD across populations, LD in Eurasian populations is somewhat larger on
average, that is, more positive for missense mutations and more negative for
loss-of-function mutations. This is in agreement with differences in
expectations between populations that have or have not gone through a
bottleneck in their recent history
(Figures~\ref{fig:toy}~and~\ref{fig:relate_dom}).

\subsubsection{Positive LD between pairs of missense mutations in conserved domains}

The similarity in signed LD between missense and synonymous mutations might
suggest that interference between missense mutations is minimal, or at least no
stronger than interference between synonymous mutations. However, interactive
effects differ dramatically between pairs of mutations found in different
intragenic regions. Due to the rarity of loss-of-function mutations, I only
compare synonymous and missense mutations when looking at finer partitions of
mutations within genes.

Annotated conserved domains in protein-coding genes drive signals of positive
LD between missense variants. Such protein-coding domains are conserved
elements of genes, often associated with some known functional or structural
feature of a protein \citep{Stanek2020-pa}. Purifying selection is expected to
be stronger within conserved domains than within the same gene but outside of
those domains.  Indeed, the site-frequency spectrum (SFS) is skewed to lower
frequencies for both missense and loss-of-function mutations within domains
when compared to the same classes of mutations outside of domains, with much
more negative values of Tajima's D within domains
(Tables~\ref{tab:tajimasD}~and~\ref{tab:msldfe}).  On the other hand, no
difference is observed for synonymous mutations whether within or outside
domains, suggesting roughly equivalent effects of selection (either direct or
linked) on synonymous variation.

Missense mutations within the same functional domain have large positive LD
that is elevated above that of synonymous mutations within the same domain
(Figures~\ref{fig:data}C~and~\ref{fig:LDwithin}--\ref{fig:domainsEAS}). This
difference between missense and synonymous variants within domains is
especially pronounced for linked pairs within a few hundred base pairs of each
other (Figure~\ref{fig:LDwithin}).

Selection is stronger against missense mutations within domains than outside
domains, leading to an excess of rare missense mutations within conserved
domains (Table~\ref{tab:tajimasD}). LD is known to be sensitive to allele
frequencies, with rare mutations showing large positive signed LD
\citep{Good2022-ot}. To test whether the signal of increased LD between
missense mutations within domains is driven by rare variants, I considered
subsets of pairs of mutations based on their derived allele frequencies
(Figures~\ref{fig:data}D--F~and~\ref{fig:domainsAFR}--\ref{fig:domainsEAS}).
Rare and uncommon variants show large average LD for each class of mutations.
However, common variants recapitulate the opposing patterns of LD that is seen
when averaging over pairs at all frequencies, and the SFS for common missense
and synonymous variants are similar (Figure~\ref{fig:sfs_proportions}), so that
this signal is unlikely to be driven by subtle differences in allele
frequencies between classes of mutations.

\subsubsection{Reduced LD between pairs of missense mutations outside of conserved domains}

The large positive signal of LD for missense mutations within the same domain
does not extend to pairs of missense mutations that span different domains.
Missense and synonymous mutations show nearly equal levels of LD close
to zero across domains, with missense mutations slightly more negative than
synonymous mutations (Figure \ref{fig:between_domains}). The interactive effect
driving large LD in domains is therefore likely domain-specific. However, the
average distance between mutations within domains is much smaller than
between domains, so this observation may be primarily driven by the higher
recombination distances between mutations across distinct domains.

Mutations that fall outside of annotated domains have the
opposite pattern of signed LD to mutations within the same
domain. For pairs of mutations outside of domains but with distances matched to
those within domains, synonymous mutations have larger positive LD
than missense mutations. More distant pairs of mutations outside of domains,
matched to the same distances as the between-domain comparison, each have LD
roughly equal to zero (Figure~\ref{fig:between_domains}).

The role that tightly linked variants have in driving these opposing signals
can be seen in the decay of signed LD with distance between mutations
(Figures~\ref{fig:LDgene}--\ref{fig:LDoutside}). Both synonymous and missense
mutation pairs at distances greater than a few hundred bases have average
LD fluctuating around zero. However, for mutations outside domains,
synonymous variants separated by short distances have large positive LD, while
missense mutations have lower LD (Figure~\ref{fig:LDoutside}). In contrast, for
mutations within the same domains, missense mutations have more positive LD at
short distances than synonymous mutations (Figure~\ref{fig:LDwithin}).

\section{Discussion}\label{sec:discussion}

%Previous theoretical and simulation studies have shown that interference and
%interactions between selected mutations reduce the efficacy of selection at
%linked loci, impacting substitution rates, the deleterious mutation load, and
%dynamics of segregating mutations
%\citep{Hill1968-vu,Birky1988-jm,Barton1995-mj,McVean2000-ox}. Interference and
%synergistic epistasis between moderately deleterious mutations are both
%expected to cause negative LD between selected mutations, which can be readily
%tested using population genetic data
%\citep{Sohail2017-zq,Sandler2021-of,Garcia2021-zn}. Here, I used a closed
%numerical approach to generate expectations for LD under a wide range of
%selective scenarios, and then compared patterns of LD in human populations
%between classes of coding mutations using unbiased estimators for LD from
%unphased genotypes.

Broad genome- and gene-wide surveys of LD across functional classes of
mutations miss heterogeneous patterns of interactive effects occurring within
genes.
From gene-wide averages, LD between missense
mutations does not appear to differ from synonymous variants,
while LD between loss-of-function variants is more negative. Hill-Robertson
effects are expected to be strongest for slightly to moderately deleterious
variants (with \(|s|\sim 1/N_e\)), as strongly deleterious mutations are not
expected to interfere with one another \citep{McVean2000-ox}. However,
inference of the distribution of fitness effects (DFE) for new loss-of-function
variants shows that a large majority are strongly deleterious (Supporting
Information), so most will not strongly interfere with each other.

Instead, negative synergistic epistasis between strongly deleterious mutations
does produce large negative deviations of mean LD. Weakly deleterious
recessive mutations can also produce this pattern, but strongly deleterious
recessive mutations lead to slightly positive LD \citep{Roze2021-cf}. While
most loss-of-function mutations are strongly deleterious, those that rise to
appreciable frequency are likely more benign and \(\sigma_d^1\) may be driven
by patterns of weakly deleterious loss-of-function mutations. The difficulty in
distinguishing these effects is compounded by the large measurement noise for
\(\E[D]\), especially for loss-of-function variants for which only a few
hundred within-gene pairs exist in the human population data analyzed here and
which are separated by larger distances on average than neighboring missense
and synonymous mutations (Figure~\ref{fig:mutDistances}).

In addition to LD varying by distance, LD can also vary due to differences in
allele frequencies among classes of mutations. Matching both distances between pairs
and allele frequencies between classes of mutations reduces these
concerns. It is further possible that recombination rates can vary between annotated
regions, resulting in differing patterns of background selection, which can
affect both allele frequencies and LD (e.g.,
Figures~\ref{fig:bgs1}--\ref{fig:bgs4}). I did not condition on local
recombination rates or inferred levels of background selection here.

\subsection{Non-uniform interactions between selected mutations within genes}

Positive average LD between both missense and synonymous mutations has been
reported in humans, \emph{Drosophila}, and other species
\citep{Sohail2017-zq,Sandler2021-of}, while others have found that
nonsynonymous mutations show lower LD than synonymous mutations
\citep{Garcia2021-zn}. The similarity of their gene-wide LD observed in this
study might suggest that interference or interactions
between missense mutations are minimal.
However, averaging over all observed pairs of
mutations within a gene masks element-specific interactive effects that drive
LD in opposite directions. Nonsynonymous mutations found within conserved
protein domains are more strongly selected
against on average, but also have increased signed LD over synonymous mutations
at the same distances within domains. Missense mutation outside of domains but
at the same distances as those within domains have more negative LD, both
compared to distance-matched synonymous mutations outside of domains and to
mutations within domains. Neither dominance effects (aside from very strongly
selected recessive mutations \citep{Roze2021-cf}), synergistic epistasis, nor
Hill-Robertson interference are expected to result in positive LD, so some
other interactive effect should be driving this signal of positive LD within
conserved domains.

Different interaction scenarios that can result in
positive signed LD between tightly linked loci. One possibility is a prevalence
of pairs of compensatory mutations that are tolerated to co-segregate at high
frequencies within conserved domains \citep{Yeang2007-gj,Ivankov2014-tn}.
\citet{Callahan2011-ac} and \citet{Taverner2020-lk} have proposed such a
mechanism to explain observed clusters of nonsynonymous substitutions in
\emph{Drosophila} and other species. Another possibility is a model of
antagonistic, or diminishing returns epistasis, in which a single amino
acid-changing mutation within a domain damages the functionality of that
subunit, but additional mutations within that same domain reduce fitness by a
factor less than the first mutation. A third possibility, related to
antagonistic epistasis, is that selection acts on the functional domain as a
unit instead of on mutations within the domain individually (such as under a
model of gene-based dominance). In this scenario, double heterozygotes have
different fitnesses depending on whether the mutations are found on same
haplotype or on different haplotypes.

There is no increase in LD between pairs of missense mutations found in
different annotated domains or outside of domains.
Rather, those missense variants have considerably
lower LD than synonymous variants, and this difference between synonymous and
missense variants disappears for SNPs separated by more than a
few hundred base pairs (Figure~\ref{fig:data}E, F). This suggests that
Hill-Robertson interference is the primary mode of interaction between missense
mutations falling outside of domains, in agreement with \citet{Garcia2021-zn},
as epistasis is expected to impact LD over larger distances than what is
observed. Importantly, however, the strength of epistasis is also likely to be
a function of distance between mutations, complicating this interpretation.

Taken together, selection and interactions in protein-coding genes are
nonuniform, depending on mutation type and location within the gene. Typical
approaches for inferring the distribution of fitness effects from population
genomic data average over such differences by aggregating all observed
nonsynonymous mutations \citep{Boyko2008-zk,Kim2017-xo}. It would be
straightforward to adapt DFE-inference methods to infer more detailed
representation of the heterogeneous effects of new mutations within genes by
partitioning by missense and loss-of-function classes as well as by annotated
domains. Additionally, standard models and simulation approaches may be too
simple to capture evolutionary trajectories and patterns of diversity that
differ due to region-specific interactions.

\subsection{Positive LD between synonymous mutations}

In a non-structured randomly mating population, neutral mutations are expected
to have average signed LD of zero, but across all populations analyzed here, LD
between synonymous mutations is positive. While selection on some subset of
synonymous variants is possible, it is likely weaker on average than
between missense mutations, and any interference between selected synonymous
variants should lead to negative LD.  Spatial population structure may be
responsible for the increase of LD observed between synonymous variants. While
neutral evolution alone in structured populations cannot cause positive LD,
\citet{Sohail2017-zq} and \citet{Sandler2021-of} used forward simulations to
show that population structure combined with selection at linked sites can
induce positive LD between neutral mutations.

Nonrandom mutational processes, in which clusters of mutations occur
simultaneously in the same mutational event, can also lead to positive LD
between neutral variants separated by short distances. Such multinucleotide
mutations have been shown to affect patterns of LD on the order of 10s to 100s
of base pairs \citep{Harris2014-zg}, and clustered mutational events may be
common in humans \citep{Besenbacher2016-ac}. Indeed, a simple exponential model
in which the fraction of mutations causing a multinucleotide mutation event
decays with distance fits the observed patterns of \(\sigma_d^1\) between
synonymous variants (Supporting Information, Figure~\ref{fig:mslmnms}), with
the best-fit model requiring only a small fraction of mutations to involve
multinucleotide mutation events.

We may therefore treat positive LD as the baseline expectation for tightly
linked variants due to either spatial structure or clustered mutations, so that
subsequent selective processes and interactions cause LD to deviate from that
expectation. It may then be more appropriate to compare the negative LD
observed between linked loss-of-function variants to that positive baseline
expectation instead of zero, which would imply that they are more recessive or
have stronger synergistic epistasis than from inferences assuming a neutral
expectation of zero. Additional analyses and simulations will be
required to tease apart the effects of population structure, selective
interference at linked sites, and clustered mutational events on patterns of
LD.

\subsection{Challenges to distinguishing modes of selective interactions from LD}

When partitioning measurements of LD by mutation classes or regions within
genes, the decreasing number of pairwise comparisons leads to large
measurement noise. Within each population, confidence intervals of observed
\(\sigma_d^1\) often overlap with zero or overlap with that of other classes of
mutations. While observed patterns are remarkably consistent across the 15
populations considered here, their joint evolutionary histories make formal
testing of significance difficult due to shared variation, as they cannot be
treated as independent measurements. Detailed simulations will likely
be needed to more thoroughly assess significance.

From a modeling perspective, the space of plausible selection scenarios becomes
large as we relax the strict assumptions of additivity and multiplicative
interactions. This makes performing forward simulations that span the range of
all such selective interaction scenarios burdensome. Instead, closed numerical
approaches allow us to efficiently explore this highly parameterized space of
models and to perform likelihood-based inference using signed LD or other
two-locus summaries of the data. For example, inferring the joint distribution
of dominance and selection is underpowered using the SFS alone, but because
signed LD is sensitive to the levels of dominance (Figure~\ref{fig:dominance}),
inferring the DFE with dominance may be feasible using the joint distribution
of allele frequencies and LD. The results presented here do not cover the space
of all possible two-locus models, and other unexplored models may result in
similar patterns of signed LD.  Comparisons to empirical data should therefore
be treated with caution.

Using a single low-order summary of signed LD, such as \(\E[D]\) or
\(\sigma_d^1\), is likely insufficient to confidently discriminate modes of
selective interactions that produce similar LD patterns. Among all interaction
models, the extent of LD and rate of its decay also depends on the underlying
distribution of selection coefficients among a class of mutations, which are
unknown for a given pair of mutations, so that we must integrate over a
distribution of fitness effects. This DFE, however, will have been inferred
under a simple set of assumptions, such as additivity and interchangeability
between sites within a gene, potentially biasing any inference using previously
inferred DFEs to learn about patterns of interactions. Again, this underlines
the need to jointly infer strengths and interactions of selected variants and
to consider patterns of variation at finer genomic scales.

Finally, expectations for a large family of informative two-locus statistics
can be computed directly from the full two-locus sampling distribution, which
can be compared to empirical observations from either phased or unphased data
\citep{Ragsdale2020-nz}. Exploring additional patterns of correlations between
mutations should uncover overlooked statistics that will improve our ability to
distinguish between modes of selective interactions.

\section{Methods}\label{sec:methods}

\subsection{Existing theory and numerical methods}

Many well-known properties of linkage disequilibrium come from
early work on the multi-locus diffusion approximation \citep{Kimura1955-qe,
Hill1968-vu, Ohta1969-ie, Ohta1971-yd}. This includes the result that
genome-wide averages of signed LD are expected to be zero under neutrality.
Under a two-locus biallelic model, where the left locus allows alleles \(A\)
and \(a\) and the right locus allows alleles \(B\) and \(b\), the standard
covariance measure of LD is defined as \(D = f_{AB} - f_{A}f_{B}\), where
\(f_{AB}\) is the haplotype frequency of types carrying both
\(A\) and \(B\), and \(f_{A}\) and \(f_{B}\) are the marginal frequencies of
those alleles at each locus. This covariance decays due to both drift and
recombination at a rate proportional to the inverse of the effective population
size and the distance separating loci \citep{Hill1968-vu}:
\[\E[D]_{t+1} = \left(1 - \frac{1}{2N_e(t)} - r \right)\E[D]_t.\]
While \(\E[D] = 0\), the variance of \(D\) is non-zero, and \citet{Ohta1971-yd}
found that the variance of \(D\) under
neutrality and steady-state demography, normalized by the joint heterozygosity
of the two loci, is
\begin{equation}
    \label{eq:ohta}
    \sigma_d^2 = \frac{\E[D^2]}{\E[f_A(1-f_A)f_B(1-f_B)]}
    \approx\frac{5 + \rho / 2}{11 + 13\rho/ 2 + \rho^2 / 2},
\end{equation}
where \(\rho = 4N_e r\).

Analytic progress beyond these results has come haltingly. In the 1980s,
recursions were developed to compute the two-locus sampling distribution under
neutrality \citep{Golding1984-pu}, and this approach later formed the
foundation for the inference of local recombination rates from population
genetic data \citep{Hudson2001-sg,McVean2004-gj}. More recently,
\citet{Song2007-qk} computed \(\E[r^2]\) using a diffusion approximation
approach, although their solution involves the summation of infinitely many
terms and is restricted to neutrality and steady-state demography.

To include selection, there have been relatively few advances beyond
the Monte Carlo simulation approach taken by \citet{Hill1966-gv}, albeit now
with more powerful computational resources and sophisticated software for
performing flexible forward simulation
\citep[e.g.,][]{Thornton2019-qc,Haller2019-vm}. Analytic results for two-locus
distributions under selection are notoriously difficult, with a few notable
flashes of progress. For example, \citet{McVean2007-xd} considered the effect
of a recent sweep on patterns of LD between neutral loci near the locus under
selection, and in a recent paper, \citet{Good2022-ot} presented analytic
solutions for patterns of LD between rare mutations under additive selection
with epistasis. Nonetheless, such approaches are typically confined to
steady-state demography and constrained selection models.

Numerical methods inhabit the space between expensive discrete simulations and
limited analytic solutions, providing a more efficient and practical method to
compute expectations of two-locus diversity measures under a wider range of
parameters and demographic scenarios. \citet{Ragsdale2017-gg} used a finite
differences approach to numerically solve the two-locus diffusion equation with
additive selection at either locus, and \citet{Ragsdale2019-nt} more recently
extended the \citet{Hill1968-vu} system for \(\E[D^2]\) to compute arbitrary
moments of the distribution of \(D\) for any number of populations connected by
gene flow. They also showed that such a moment-construction can be used to
solve for the two-locus sampling distribution within a single population,
though it requires a moment-closure approximation for nonzero recombination and
selection. Below, I extend this approach to model arbitrary diploid selection,
which encompasses dominance, epistasis, and other forms of selective
interactions between two loci. In a concurrent study to this paper,
\citet{Friedlander2022-bs} developed numerical solutions to the same moment
system, which they used to describe selected haplotype trajectories and the
distortion of neutral diversity at loci variably linked to beneficial alleles
that sweep to high frequencies under non-equilibrium demography.

\subsection{The two-locus sampling distribution with arbitrary selection}

The two-locus sampling distribution is the direct analog to the single-locus
site frequency spectrum (SFS) of a given sample size
(Figure~\ref{fig:TwoLocusFS}). The two-locus
distribution \(\Psi_n\) stores the density or number of pairs of loci with
observed haplotype counts in a sample of size \(n\),
so that \(\Psi_n(i, j, k)\) is the number of pairs
for which we observe \(i\) copies of the \(AB\) haplotype, \(j\) of type
\(Ab\), \(k\) of type \(aB\), and \(n-i-j-k\) of type \(ab\). The size of
\(\Psi_n\) grows rapidly, with \(O(n^3)\) entries, which practically limits
computational approaches to moderate sample sizes \(n\lesssim 100\) and a
single population.

Under neutrality, many approaches exist to compute \(\Psi_n\), including the
recursion due to \citet{Golding1984-pu} and \citet{Ethier1990-tb} and recent
numerical approaches \citep{Kamm2016-ag, Ragsdale2017-gg}. Selection is most
easily included using the forward-in-time diffusion equation
\citep{Kimura1955-qe,Hill1966-gv}, where a standard approach is to first solve
for the continuous distribution \(\psi\) of the density of two-locus haplotype
configurations in the full population, and then integrate \(\psi\) against the
multinomial sampling function to obtain \(\Psi_n\). Alternatively,
\citet{Ragsdale2019-nt} showed that there exists a system of ordinary
differential equations directly on the entries of \(\Psi_n\). I briefly
summarize this general approach below, but refer readers to that paper for
detailed derivations of the drift, recombination, and mutation terms and the
moment-closure approximation. Instead, here I focus on generalizing the
selection operator to include epistasis, dominance, and other forms of
two-locus interactions.

\subsubsection{Moment equation for \(\Psi_n\)}

The system of linear ordinary differential equations for the entries of
\(\Psi_n\) takes the form
\begin{equation}
\label{eq:system}
{\Psi}_n^{t+1}(i, j, k; t) =
\mathcal{D}_{N(t)}\Psi_n^t
+ \mathcal{R}_{r}\Psi_{n+1}^t
+ \mathcal{U}_{u}\Psi_n^t
+ \mathcal{S}_{s,h}\Psi_{n+2}^t.
\end{equation}
\(\mathcal{D}_{N(t)}\) is a sparse linear operator accounting for drift
with population size \(N(t)\), \(\mathcal{R}\) accounts for recombination
with per-generation recombination probability \(r\) between loci,
\(\mathcal{U}\) accounts for mutation, either under an infinite sites or
biallelic reversible mutation model, and \(\mathcal{S}\) accounts for
selection.

The moment system for \(\Psi_n\) can be derived directly from the diffusion
approximation, or it can be found through a more intuitive process of tracking
the dynamics of allelic states of a sample of size of \(n\) from the full
population. We assume \(n \ll N_e\), and \(r\) and \(s\) are \(O(1/N_e)\) so
that multiple coalescence, recombination, or selective events within the \(n\)
lineages are rare in any given generation \citep[Supporting
Information;][]{Jouganous2017-pq,Ragsdale2019-nt}. In typical diffusion
approximation fashion, we multiple through by \(2N_{ref}\) so that time is
measured in \(2N_{ref}\) generations, and we consider scaled parameters \(\rho
= 4Nr\), \(\theta = 4Nu\), and \(\gamma=2Ns\).

\subsubsection{Moment closure}\label{moment-closure}

In the absence of selection and for fully linked loci (\(\rho=0\)), the system
closes and can be solved exactly. However, for nonzero recombination or
selection, the entries of \(\Psi_n\) rely on the slightly larger sampling
distributions with sample sizes \(n+1\) (for recombination and additive
selection) or \(n+2\) (for non-additive selection). This is because if a
recombination event occurs within one of \(n\) lineages being tracked by
\(\Psi_n\), we need to draw an additional lineage from the full population to
recombine with that chosen lineage, thus requiring \(\Psi_{n+1}^t\) to find
\(\Psi_n^{t+1}\). Selection events similarly require extra lineages from the
full population, which replace a chosen lineage that fails to reproduce with
probability proportional to its relative fitness.

This requirement of extra lineages for recombination and selection means that
the system in \eqref{eq:system} is not closed, so that we need a moment-closure
approximation to solve for \(\Psi_n\). As in \citet{Ragsdale2019-nt}, a
jackknife approximation is used to estimate \(\Psi_{n+l}\), for \(l=1\) or
\(2\), from \(\Psi_n\) \citep[following the single-locus closure introduced
in][]{Jouganous2017-pq}, so that \(\hat{\Psi}_{n+l}(i, j, k) = \mathcal{J}_{n,
l}\Psi_n\), although other accurate closure approximations are possible
\citep{Friedlander2022-bs}. This emits a closed approximate system,
\begin{equation}
\label{eq:closed-system}
\dot{\hat\Psi}_n(i, j, k; t) =
\mathcal{D}_{\nu(t)}\hat\Psi_n(t)
+ \mathcal{R}_{\rho}\mathcal{J}_{n, 1}\hat\Psi_{n}(t)
+ \mathcal{U}_{\theta}\hat\Psi_n(t)
+ \mathcal{S}_{\gamma, h}\mathcal{J}_{n, 2}\hat\Psi_{n}(t).
\end{equation}

The jackknife approximation, which approximates an entry \(\Psi_{n+l}(i,j,k)\)
using nearby entries in \(\Psi_n\), is more accurate for larger sample sizes,
creating a tension between efficiency and accuracy: larger sample sizes result
in more accurate solutions, as error in the jackknife is diminished, but
computational complexity also grows rapidly in the number of entries of
\(\Psi_n\), which is \(O(n^3)\) (Figure~\ref{fig:jackknife}). In the results
presented in this paper, sample sizes between \(n=30\) and \(n=80\) are used.
Derivations for the drift, recombination, and mutation operators and the
jackknife moment-closure approximation can be found in section S1.3 of
\citet{Ragsdale2019-nt}, and I repeat the main results in the Supporting
Information of this paper.

\subsubsection{Selection models with epistasis and dominance}

To include selection, we consider a model where we draw lineages uniformly from
the previous generation, but keep lineages with probability proportional to
their fitness. In the absence of dominance, selection reduces to a haploid
model, with acceptance and rejection probabilities depending on the fitnesses
of each haploid copy, where haplotype \(Ab\) has fitness \(1 + s_{A}\), \(aB\)
has fitness \(1 + s_{B}\), and \(AB\) has fitness \(1 + s_{AB}\). We assume the
doubly ancestral haplotype \(ab\) has fitness \(1\), so fitnesses are relative
to that of \(ab\) haplotypes. The standard multiplicative fitness function
assumes that \(s_{AB} \approx s_{A} + s_{B}\) (assuming \(s^2\approx0\)), and a
model for epistasis can be written as \[s_{AB} = (s_{A} + s_{B}) (1 +
\epsilon),\] so that \(\epsilon > 0\) gives synergistic epistasis and
\(\epsilon < 0\) gives antagonistic epistasis.

To obtain the recursion equation under selection we consider drawing \(n\)
lineages from generation \(t\), which has an expected sampling distribution of
haplotype counts given by \(\Psi_n^t\). However, assuming \(s\leq0\) for each
derived haplotype, each of those sampled lineages has probability of being
rejected equal to the absolute value of the selection coefficient assigned to
its haplotype state. If a lineage is rejected, a replacement is drawn from the
full population. Under the assumption that \(ns \ll 1\), the probability that
more than one selection event occurs in any given generation is negligibly
small, so that the case of multiple simultaneous rejections can be ignored.
Then \(\Psi_n^{t+1}\) relies only on \(\Psi_n^t\) and \(\Psi_{n+1}^t\) for
additive selection. The full selection operator \(\mathcal{S}\) for additive
selection is given in the Supporting Information.

To account for dominance, or other general forms of two-locus selection, the
selection operator no longer reduces to individual haplotypes, but instead we
need to know the state of two-locus genotypes. For example, the fitness of an
individual carrying an \(Ab\) haplotype depends on whether their second
haplotype is \(ab\), \(Ab\), \(aB\), or \(AB\). We can therefore assign a
selection coefficient to each possible diploid configuration, \(s_{Ab/ab}\),
\(s_{Ab/Ab}\), and so on. Assuming that the doubly homozygous ancestral
\(ab/ab\) genotype has relative fitness 1, this gives nine possible unique
selection coefficients in the most general two-locus selection model. Doubly
heterozygous \(AB/ab\) and \(Ab/aB\) genotypes need not have the same selection
coefficient (Table \ref{tab:selmodels}).

The general selection operator follows the same approach as the haploid
selection operator with epistasis described above. Now, in the case of a
selection event rejecting a lineage within our tracked samples, we need to draw
not only the replacement lineage from the full population but also a second
haplotype from the full population to form the diploid genotype, as this
determines the probability that we reject the focal haplotype. We again assume
that \(ns \ll 1\) for all genotype selection coefficients, so that we may
assume at most a single selection event occurs in any given generation. This
means that to find \(\Psi_n^{t+1}\) under a general two-locus selection model
we need \(\Psi_{n+2}^t\). Again, a full derivation and expressions for the
general selection operator are given in the Supporting Information.

\subsubsection{Low-order summaries of the sampling distribution}

From \(\Psi_n\), expectations for any two-locus statistic can be found by
downsampling to the appropriate sample size. For example, to compute \(\E[D]\),
the sum is taken over all haplotype configurations \(\mathbf{n} = (n_{AB},
n_{Ab}, n_{aB}, n_{ab})\), weighted by the density \(\Psi_n\) for that
configuration:
\begin{equation}
\E[D] = \sum_{\mathbf{n}} \Psi_n(\mathbf{n})
\frac{n_{AB}n_{ab} - n_{Ab}n_{aB}}{n(n-1)}.
\end{equation}
For large sample sizes, this is approximately equal to computing \(D\) by
taking the maximum likelihood estimate for each allele frequency \(f_i = n_i /
n\), but the maximum likelihood-based estimate will be noticeably biased for
small to moderate sample sizes. Other low-order two-locus statistics can be
computed using the same approach, as implemented in \texttt{moments} following
\citet{Ragsdale2020-nz}, which can be compared across sample sizes and between
estimates from phased or unphased data. In this paper, I focus on \(\sigma_d^2
= \E[D^2] / \E[p(1-p)q(1-q)]\) and \(\sigma_d^1 = \E[D] / \E[p(1-p)q(1-q)]\),
which can be averaged over pairs of variants at all frequencies.
Allele-frequency conditioned statistics (such as keeping only loci below some
frequency threshold as in \citet{Good2022-ot}) can be considered using this
same approach.

\subsubsection{Simulations of non-steady-state demography}

I consider four variable population size histories, two simple toy models and
two inferred from human populations in African and Europe using \texttt{Relate}
\citep{Speidel2019-nj}. For each size history scenario, I track the evolution
of \(\Psi_n(t)\) for varying selection models, plotting the trajectories of
\(\sigma_d^1\) and \(\sigma_d^2\) over time (Figures~\ref{fig:toy},
\ref{fig:relate_dom}, and \ref{fig:toy_sd2}--\ref{fig:relate_dom_sd2}). The
selection strength at both loci is fixed at either \(\gamma=-1\) or \(-10\)
for the models with epistasis, or \(\gamma=-2\) for models with dominance, and
recombination is set to zero.

The simple size change models both have ancestral \(N_e=10,000\), with one a
\(3\)-fold population expansion that occurs 3,000 generations ago, and the
other a \(5\)-fold reduction 2,000 generations ago followed by a recovery to
its initial size \(1,000\) generations ago. The size histories for YRI and CEU
are inferred using \texttt{Relate} \citep{Speidel2019-nj} applied to the phase
3 haplotype-phased autosomal data from
\citet{1000_Genomes_Project_Consortium2015-zq}, using default parameters as
recommended in the \texttt{Relate} online tutorial, assuming a mutation rate of
\(1.25\times10^{-8}\) per-bp per-meiosis and a human generation time of 29
years. \texttt{Relate} returns estimates of coalescence rates within specified
time bins, and population sizes are estimated as their inverses. Estimates
using \texttt{Relate} for population sizes in the very recent past (\(<3,000\)
years, or \(\approx 100\) generations) diverge, so I truncate the history
over this time period and assume a constant size from the most recent
non-diverged bin.

\subsection{Analysis of human genomic data}
\label{analysis-of-human-genomic-data}

Using the annotated variant call format (VCF) files from the phase 3
\citet{1000_Genomes_Project_Consortium2015-zq} (Thousand Genomes) data release,
I subset the genotype VCFs to autosomal variants that are annotated as either
synonymous or nonsynonymous, including both missense mutations and more
damaging ``high impact'' loss-of-function mutations. Loss-of-function
annotations include frameshifts, splice acceptor, splice donor, start loss,
stop gain, stop loss, and transcript ablation variants. I further subset to
samples within each non-admixed population in the African, European, and East
Asian continental groups (five populations each, Table \ref{tab:1kgpops}).
Signed LD is sensitive to ancestral state misidentification, so I only keep
sites for which ancestral alleles were estimated with high confidence in both
the VCF info field and the Thousand Genomes human ancestor reconstructed from a
phylogeny of six primates.

In addition to ancestral state misidentification, measured LD is sensitive to
phasing error, so I compute LD statistics using unphased genotypes following
\citet{Ragsdale2020-nz}. This approach provides unbiased estimates for pairwise
LD, under the assumption that individuals are not inbred. I consider pairs of
mutations within the same mutation class (synonymous, missense, and
loss-of-function) either within the same gene and inside or outside of
annotated domains within the same protein-coding genes. I use a dataset of
annotated protein domains mapped to the hg19 human reference build compiled by
\citet{Stanek2020-pa} to determine if a given mutation falls within our outside
a conserved domain.

\subsection{Data and software availability}
\label{data-and-software-availability}

All data and software used in this paper are publicly available and open
source. I downloaded the Thousand Genomes annotations and genotypes VCFs from
the ftp server at
\url{ftp://ftp.1000genomes.ebi.ac.uk/vol1/ftp/release/20130502/}, and the
Thousand Genomes human ancestor fasta file from
\url{ftp://ftp.1000genomes.ebi.ac.uk/vol1/ftp/phase1/analysis_results/supporting/ancestral_alignments/}.
Protein domain information from \citet{Stanek2020-pa} was downloaded from
\url{http://prot2hg.com/dbdownload.php}.

Implementation of moment equations to compute expectations for two-locus and
linkage disequilibrium statistics are implemented in Python using Numpy
\citep{Harris2020-pc} and sparse matrix solvers in Scipy
\citep{Virtanen2020-kr}. These methods are packaged within \texttt{moments},
and analyses here were performed using \texttt{moments} version 1.1.10,
available from \url{https://bitbucket.org/simongravel/moments/src/master/} and
via \texttt{conda}, with extensive documentation at
\url{https://moments.readthedocs.org}. Scripts to run all analyses, recreate
figures, and compile this manuscript are available at
\url{https://github.com/apragsdale/two_locus_selection}.

\section{Acknowledgments}\label{acknowledgements}

I thank Alex Diaz-Papkovich, Eric Friedlander, Simon Gravel, Mashaal Sohail,
Matthias Steinrücken, and Kevin Thornton for helpful discussions and feedback
on earlier versions of this manuscript.

\bibliographystyle{genetics}  
\bibliography{paper}

\end{document}